### 2020-04-20 Manuscript/SupportingInfo.Rmd

---
title:

output: 
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: true
    pandoc_args:

csl: evolution.csl
bibliography: Lambda-Refs.bib

---
# Supporting Information for: Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate

**FOR DCA**

1: balanced/symmetric trees
2: Kappa and K_Z
3: Beta from PGLS

Here we provide additional supporting information referenced in the main document: additional analyses, and simulation results across a wider set of input conditions.

## Simulations on differently shaped phylogenies

```{r setup, include=FALSE}
  library(knitr)
  knitr::opts_chunk$set(echo = TRUE)
```

In addition to using pure-birth phylogenies, we explored the effect of tree shape on our findings using both balanced and pectinate trees. As before, simulations were conducted on a range of tree sizes ($n=2^5 - 2^{10}$), and across a range of input levels of phylogenetic signal ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units). For each $n$ and $\lambda_{in}$ combination, 50 replicates of a continuous trait were simulated using a Brownian motion model of evolution. Using these, we estimated the degree of phylogenetic signal using $\lambda$. \hfill\break

*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied ..... ** FINISH THIS** (Fig. S1, S2).

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig1.png")

```

\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using balanced trees. \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("Fig1.png")

```

\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees. \hfill\break

## Literature Survey

\newpage

# References

 \setlength{\parindent}{-0.25in}
 \setlength{\leftskip}{0.25in}
 \setlength{\parskip}{8pt}
 \noindent

<div id="refs"></div>