### 2020-05-05 Manuscript/SupportingInfo-EKB.Rmd
Draft of Supporting Information ready for edits. All robustness analyses have been finished, and corresponding supporting figures have been generated. One has to copy said figures over to manuscript folder in order to knit SupportingInfo-EKB.Rmd.

---
title:

output: 
  pdf_document:
    fig_caption: yes
    number_sections: no  #note: change to 'no' for most journals
    keep_tex: true
    pandoc_args:

csl: evolution.csl
bibliography: Lambda-Refs.bib

---
# Supporting Information for: Estimates of Phylogenetic Signal Based on Lambda are Often Inaccurate

*COPY FIGURES FROM FIGURES FOLDER INTO MANUSCRIPT FOLDER IN ORDER TO KNIT*

Here we provide additional supporting information referenced in the main document: additional analyses, and simulation results across a wider set of input conditions.

## Simulations on differently shaped phylogenies

```{r setup, include=FALSE}
  library(knitr)
  knitr::opts_chunk$set(echo = TRUE)
```

In addition to using pure-birth phylogenies, we explored the effect of tree shape on our findings using both balanced and pectinate trees. As before, simulations were conducted on a range of tree sizes ($n=2^5 - 2^{10}$), and across a range of input levels of phylogenetic signal ($\lambda_{in} = 0.0 \to 1.0$; in 21 intervals of 0.05 units). For each $n$ and $\lambda_{in}$ combination, 50 replicates of a continuous trait were simulated using a Brownian motion model of evolution. Using these, we estimated the degree of phylogenetic signal using $\lambda$. \hfill\break

*Results*. As found with pure-birth phylogenies, estimates of $\lambda$ varied most dramatically in simulations with fewer species and at intermediate values of lambda (Fig. S1, S2). Pectinate trees showed an interesting tendency to underestimate $/lambda$ as 0 for input $/lambda$ values. Most dramatically, some simulations on the largest pectinate phylogeny ($n = 1024$) estimated $/lambda = 0$ for input values as high as 0.75 (Fig. S2).

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS1.png")

```

\textbf{Figure S1}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using balanced trees. \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS2.png")

```

\textbf{Figure S2}. Accuracy of Pagel's lambda estimations across known lambda inputs on various tree sizes. Results obtained using pectinate trees. \hfill\break

\newpage

## Simulations of phylogenetic regression and ANOVA

We analyzed the reliability of parameter estimations when used in phylogenetic regression and ANOVA. This involved simulations of dependent variables with $\lambda$ values ranging from 0 to 1 (21 intervals of 0.05 units) across pure-birth trees ($n=2^5 - 2^{10}$). For each $\lambda$ input value, independent variables were then generated with a known relationship to the simulated dependent variable ($\beta = 0, 0.25, 0.5, 0.75, 1.0$), from which $\lambda$ was then estimated.

*Results* As with the results from the regression analyses in the main text, we found poor precision of $\lambda$ estimation in smaller trees and at intermediate levels of phylogenetic signal with phylogenetic ANOVA (Fig. S3). Slope was reliably estimated across tree sizes and known $\lambda$ values (Fig. S4).

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS3.png")

```

\textbf{Figure S3}. Precision of Pagel's $\lambda$ when incorporated in phylogenetic ANOVA. Results are from input $\beta$ values of $0.5$. \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS4.png")

```

\textbf{Figure S4}. Precision of slope estimates across known input $\beta$ values of phylogenetic regression. Results shown include all input $\lambda$ values. \hfill\break

\newpage

## Effect Sizes of Pagel's Lambda and Kappa

In this manuscript, we propose the use of effect size (Z-scores) over raw metrics of phylogenetic signal. We quantified Z-scores for estimated values of Pagel's $\lambda$ and Kappa of simulated datasets to demonstrate the variable behavior of these effect sizes across tree size and input values of phylogenetic signal. Simulation procedures can be found in the main text.

*Results* Z scores of Pagel's $\lambda$ show a dramatically non-linear pattern across values of known $\lambda$ across all tree sizes (Figure S5). On the other hand, variation in the Kappa Z-scores is consistent across input $\lambda$s and shows a nearly linear relationship. Furthermore the variation of $Z_K$ is fairly consistent across values of input $\lambda$ indicating that $Z_K$ is an accurate and reliable statistic and can be appropriately used when comparing values of phylogenetic signal across clades or traits.

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS5.png")

```

\textbf{Figure S5}. Variation in effect size (Z score) of Pagel's $\lambda$ across input $\lambda$ values on various tree sizes.  \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS6.png")

```

\textbf{Figure S6}. Variation in effect size (Z score) of $K$ across input $\lambda$ values on various tree sizes. \hfill\break

```{r, echo = FALSE, out.width="95%"}
include_graphics("FigS7.png")

```

\textbf{Figure S7}. Accuracy of Kappa estimations across known lambda inputs on various tree sizes.  \hfill\break

\newpage

# References

 \setlength{\parindent}{-0.25in}
 \setlength{\leftskip}{0.25in}
 \setlength{\parskip}{8pt}
 \noindent

<div id="refs"></div>