### 2020-07-12 Manuscript/Manuscript-7-MLC.tex
version 8 started

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\hypertarget{a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}{%
\section{A Standardized Effect Size for Evaluating the Strength of
Phylogenetic Signal, and Why Lambda is not
Appropriate}\label{a-standardized-effect-size-for-evaluating-the-strength-of-phylogenetic-signal-and-why-lambda-is-not-appropriate}}

\hfill\break

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, and wish to compare the strength of that signal
across traits. However, analytical tools for such comparisons have
largely remained underdeveloped. In this study, we evaluated the
efficacy of one commonly used parameter (Pagel's \(\lambda\)) to
estimate the strength of phylogenetic signal in phenotypic traits, and
evaluate the degree to which \(\lambda\) correctly identifies known
levels of phylogenetic signal. We find that the precision of \(\lambda\)
in estimating actual levels of phylogenetic signal is often inaccurate,
and that biological interpretations of the strength of phylogenetic
signal based on \(\lambda\) are therefore compromised. We then propose a
standardized effect size based on \(\kappa\) (\(Z_\kappa\)), which
measures the strength of phylogenetic signal, and places it on a common
scale for statistical comparison. Tests based on \(Z_\kappa\) provide a
mechanism for formally comparing the strength of phylogenetic signal
across datasets, in much the same manner as effect sizes may be used to
summarize patterns in quantitative meta-analysis. Our approach extends
the phylogenetic comparative toolkit to address hypotheses that compare
the strength of phylogenetic signal between various phenotypic traits,
even when those traits are found in different evolutionary lineages or
have different units or scales.

\newpage

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry among species
violates an assumption of independence among trait values that is common
for statistical tests (Felsenstein 1985; Harvey and Pagel 1991).
Accounting for this evolutionary non-independence is the purview of
\emph{phylogenetic comparative methods} (PCMs): a suite of analytical
tools that condition trends in the data on the phylogenetic relatedness
of observations (e.g., Grafen 1989; Garland and Ives 2000; Rohlf 2001;
Butler and King 2004). The past several decades have witnessed a rapid
expansion in the development of PCMs to address an ever-growing set of
macroevolutionary hypotheses (Martins and Hansen 1997; O'Meara et al.
2006; Revell and Harmon 2008; Beaulieu et al. 2012; Adams 2014b,a; Adams
and Collyer 2018). These methods are predicated on the notion that
phylogenetic signal -- the tendancy for closely related species to
display similar trait values -- is present in cross-species datasets
(Felsenstein 1985; Pagel 1999; Blomberg et al. 2003). Indeed, under
numerous evolutionary models, phylogenetic signal is to be expected, as
stochastic character change along the hierarchical structure of the tree
of life generates trait covaration among related taxa (see Felsenstein
1985; Blomberg et al. 2003; Revell et al. 2008). \hfill\break

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets, including measures of serial independence
(\(\mathbf{C}\): Abouheif 1999), autocorrelation estimates (\(I\):
Gittleman and Kot 1990), statistical ratios of trait variation relative
to what is expected given the phylogeny (\(\kappa\): Blomberg et al.
2003; Adams 2014a), and scaling parameters used in maximum likelihood
fitting of the data to the phylogeny (\(\lambda\): Pagel 1999), among
others (e.g., Klingenberg and Gidaszewski 2010). The statistical
properties of these methods -- namely type I error rates and power --
have also been investigated to determine when phylogenetic signal can be
detected and under what conditions (e.g., Münkemüller et al. 2012;
Pavoine and Ricotta 2012; Diniz-Filho et al. 2012; Adams 2014a;
Molina-Venegas and Rodríguez 2017; see also Revell et al. 2008; Revell
2010). One of the most widely used methods for characterizing
phylogenetic signal in macroevolutionary studies is Pagel's \(\lambda\)
(Pagel 1999). The parameter (\(\lambda\)) transforms the lengths of the
internal branches of the phylogeny to improve the fit of data to the
phylogeny via maximum likelihood (Pagel 1999; Freckleton et al. 2002).
Pagel's \(\lambda\) ranges from \(0\to1\), with larger values signifying
a greater dependence of observed trait variation on the phylogeny.
Pagel's \(\lambda\) also has the appeal that it may be included in
phylogenetic generalized least-squares regression (PGLS) to account for
the degree of phylogenetic signal in comparative analyses (see
Freckleton et al. 2002). \hfill\break

In PGLS analysis, \(\lambda\) is a parameter that is tuned via
log-likelihood profiling for its estimation. After obtaining an
optimized \(\lambda\) value, \(\hat{\lambda}\), likelihood ratio tests
can be performed to compare observed model fits using \(\hat{\lambda}\)
to those obtained when \(\lambda=0\) or \(\lambda=1\) to infer whether
phylogenetic signal differs from no signal or a Brownian motion (BM)
model of evolutionary divergence, respectively (Freckleton et al. 2002;
Cooper et al. 2010; Bose et al. 2019). Confidence limits on
\(\hat{\lambda}\) can be also estimated based on percentiles of a
\(\chi^2\) distribution and where these values are mapped onto a
log-likelihood profile (\emph{sensu} Freckleton et al. 2002). With
respect to confidence limits, one can ascertain whether \(\lambda=0\) or
\(\lambda=1\) is contained within an interval of \(\hat{\lambda}\), as
an analog to likelihood ratio tests (see, e.g., Vandelook et al. 2019).
It is, therefore, tempting to regard \(\hat{\lambda}\) as a descriptive
statistic that measures the relative strength of phylogenetic signal,
for inferring the extent to which shared evolutionary history has
influenced trait covariation among taxa, is common. The appeal of
\(\hat{\lambda}\) as a descriptive statistic is that it serves as the
basis for interpreting ``weak'' versus ``strong'' phylogenetic signal;
i.e., small versus large values of \(\hat{\lambda}\), respectively, in a
comparative sense (e.g., De Meester et al. 2019; Pintanel et al. 2019;
Su et al. 2019). For example, qualitative comparisons of
\(\hat{\lambda}\) for multiple traits on the same phylogenetic tree
provide a basis for inferring whether the strength of phylogenetic
signal is greater in one trait as compared to another (e.g., Liu et al.
2019; Bai et al. 2019). Indeed, statements regarding the strength of
phylogenetic signal based on \(\hat{\lambda}\) are rather common in the
evolutionary literature. For instance, of the 204 papers published in
2019 that estimated and reported Pagel's \(\lambda\) (found from a
literature survey we conducted in Google.scholar), 40\% explicitly
interpreted the strength of phylogenetic signal for at least one
phenotypic trait. Further, because nearly half of the 1,572
\(\hat{\lambda}\) values reported were near 0 or 1 (Figure 1) where the
biological interpretation of \(\lambda\) is known (\(\lambda=0\)
represents no phylogenetic signal, while \(\lambda=1\) is phylogenetic
signal as expected under BM), this percentage is even higher.
\hfill\break

{[}insert Figure 1 here{]} \hfill\break

It seems intuitive to interpret the strength of phylogenetic signal
based on the value of \(\hat{\lambda}\), as \(\lambda\) is a parameter
on a bounded scale (\(0\to1\)) for which interpretation of its extremal
points are understood. However, equating values of \(\hat{\lambda}\)
directly to the strength of phylogenetic signal presumes two important
statistical properties that have not been fully explored. First, it
presumes that values of \(\hat{\lambda}\) can be precisely estimated, as
biological inferences regarding the strength of phylogenetic signal
depend on high accuracy in its estimation. (In order to be precise, the
log-likelihood profile would have to be peaked rather than flat.)
Therefore, understanding the precision in estimating \(\hat{\lambda}\)
is paramount. One study (Boettiger et al. 2012) found that estimates of
Pagel's \(\lambda\) displayed less variation (i.e., greater precision)
when data were simulated on a large phylogeny (\(N=281\)) as compared to
a small one (\(N=13\)). From this observation it was concluded that
insufficient data (i.e., low number of species) was the underlying cause
of the increased variation across parameter estimates (Boettiger et al.
2012). Indeed, such a pattern is common with statistical estimators, as
summary statistics and parameters are often more precise at greater
sample sizes (Cohen 1988). However, this conclusion could imply that the
shape of log-likelihood profiles are assumed consistent with species
number. Such an assumption and the importance of tree shape have to date
not been verified. Thus, despite widespread use of Pagel's (1999)
\(\lambda\) in macroevolutionary studies, at present, we lack a general
understanding of the precision with which \(\hat{\lambda}\) can estimate
levels of phylogenetic signal in phenotypic datasets. \hfill\break

Second, while estimates of \(\lambda\) are within a bounded scale
(\(0\to1\)), this does not \emph{de-facto} imply that the estimated
values, \(\hat{\lambda}\), of this parameter correspond to the actual
strength of the underlying input signal in the data. For this to be the
case, \(\hat{\lambda}\) must be a statistical effect size. Effect sizes
are a measure of the magnitude of a statistical effect in data,
represented on a common scale (Glass 1976; Cohen 1988). Effect sizes
have widespread use in many areas of the quantiative sciences, as they
represent measures that may be readily summarized across datasets as in
meta-analyses (Glass 1976; Hedges and Olkin 1985; Arnqvist and Wooster
1995), or compared among datasets (e.g., Adams and Collyer 2016, 2019a).
Unfortunately, not all model parameters and descriptive statistics are
effect sizes, directly, and thus many summary measures must first be
converted to statistics with standardized units (i.e., conversion to an
effect size) for meaningful comparison (see Rosenthal 1994). As a
consequence, it follows that only if \(\hat{\lambda}\) is a statistical
effect size can comparisons of estimates across datasets be
interpretable. However, the calculation and statistical behavior of
\(\hat{\lambda}\) as an effect size has not yet been explored.
\hfill\break

As a prortional random variable bounded by 0 and 1, we might expect that
\(\hat{\lambda}\) follows A Bernoulli distribution; i.e., branch lengths
in a tree are scaled proportionally to the probability that data arise
from a BM process. As such, we would also expect in simulation
experiment that given a known \(\lambda\) value used to generate random
data on a tree: the mean of an empirical sampling distribution of
\(\hat{\lambda}\) would approximately equal \(\lambda\); the standard
error of \(\hat{\lambda}\) would be largest at intermediate values of
\(\lambda\), and predictable over the range of \(\lambda\) with respect
to treesize; the distribution of \(\hat{\lambda}\) would be symmetric at
intermediate values of \(\lambda\) and more skewed toward values of 0 or
1; and that the distribution of \(\hat{\lambda}\) will be more
platykurtic at intermediate values of \(\lambda\), becoming more
leptokurtic toward 0 and 1. These properties seemed to be
superficiallyreasonable from the simulations perfomed by Münkemüller et
al.(2012, see Fig. 2). Their study simulated strength of phylogenetic
signal as a varied weighted-average of data simulated on trees with
\(\lambda=0\) and \(\lambda=1\), rather than from prescribed values of
\(\lambda\). Nonetheless, their results revealed that as a random
variable, \(\hat{\lambda}\) resembled a Bernoulli random variable based
superficially on statistical moments (mean, variance, skewness, and
kurtosis; see their Fig. 2) with respect to strength of phylogenetic
signal, for a given tree size. However, for comparisons among tree sizes
for a given strength of phylogenetic signal, median values of
\(\hat{\lambda}\) were slightly, positively associated with the number
of species in a tree, especially at weaker strength of signals,
suggesting \(\hat{\lambda}\) might not be an accurate statistic across a
range of tree sizes. In order for \(\hat{\lambda}\) to be considered a
reliable statistic as a measure of phylogenetic signal, as a minimum, it
should be accurate and precise, properties that could be measured from
simulation experiments for known values of \(\lambda\).

In this study, we evaluate the precision of \(\hat{\lambda}\) as an
optimized parameter, used as a statistic, for estimating known levels of
phylogenetic signal in phenotypic data (data simulated with known
\(\lambda\)). We use computer simulations with differing numbers of
species, differently shaped phylogenies, and differing input levels of
\(\lambda\), to explore the degree to which \(\hat{\lambda}\) correctly
identifies known levels of phylogenetic signal, \(\lambda\), and under
which circumstances. We find that estimates of \(\hat{\lambda}\) vary
widely for a given input value of \(\lambda\), and that the precision of
\(\hat{\lambda}\) is not constant across its range. Rather, there is
decreased precision when input levels of phylogenetic signal are of
intermediate strength. Additionally, the same values of
\(\hat{\lambda}\) may be obtained from datasets containing vastly
different input levels of \(\lambda\). Thus, \(\hat{\lambda}\) is not a
reliable indicator of the strength of phylogenetic signal in phenotypic
data. We then describe a standardized effect size for measuring the
strength of phylogenetic signal in phenotypic datasets and apply the
concept to two common measures of phylogenetic signal: \(\hat{\lambda}\)
and \(\kappa\). Through simulations we find that the precision of effect
sizes based on \(\hat{\lambda}\) (\(Z_{\lambda}\)) are less reliable
than that those based on \(\kappa\) (\(Z_\kappa\)), implying that
\(Z_\kappa\) is a more robust effect size measure. We also propose a
two-sample test statistic that may be used to compare the strength of
phylogenetic signal among datasets, and provide an empirical example to
demonstrate its use. We conclude that estimates of phylogenetic signal
using Pagel's \(\lambda\) (as a descriptive statistic, not as a PGLS
parameter) are often inaccurate, and thus interpreting strength of
phylogenetic signal in phenotypic datasets based on this measure is
compromised. By contrast, effect sizes obtained from \(\kappa\) hold
promise for characterizing phylogenetic signal, and for comparing the
strength of phylogenetic signal across datasets.

\hypertarget{methods-and-results}{%
\section{Methods and Results}\label{methods-and-results}}

\hypertarget{the-precision-of-hatlambda-is-variable}{%
\subsection{\texorpdfstring{\emph{The Precision of \(\hat{\lambda}\) is
Variable}}{The Precision of \textbackslash{}hat\{\textbackslash{}lambda\} is Variable}}\label{the-precision-of-hatlambda-is-variable}}

We conducted a series of computer simulations to evaluate the precision
of \(\hat{\lambda}\) based on varied input of Pagel's \(\lambda\) to
generate data. Our primary simulations were based on pure-birth
phylogenies; however, we also evaluated patterns on both balanced and
pectinate trees to determine whether tree shape affected our findings
(see Supporting Information). First we generated 50 pure-birth
phylogenies at each of six different tree sizes, ranging from 32 to 1024
taxa (\(n=2^5 - 2^{10}\)). For each simulation run, we rescaled the
simulated phylogenies by multiplying the internal branches by
\(\lambda_{in}\), using 21 intervals of 0.05 units across its range
(\(\lambda_{in} = 0.0 \to 1.0\)), resulting in 1050 scaled phylogenies
at each level of species richness (\(n\)). Continuous traits were then
simulated on each phylogeny under a BM model of evolution to obtain
datasets with differing levels of phylogenetic signal, that ranged from
no phylogenetic signal (when \(\lambda_{in} =0\)), to phylogenetic
signal reflecting a BM model of evolution (when \(\lambda_{in} =1\)).
For each dataset we then estimated phylogenetic signal
(\(\hat{\lambda}\)), and calculated its variance
(\(\sigma^2_{\hat{\lambda}}\)) at each input level of phylogenetic
signal and level of species richness as an estimate of precision. Random
data samples were generated from a normal distribution, i.e.,
\(\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\). We verified that the
level of \(\sigma^2\) had no effect on \(\hat{\lambda}\) or
\(\sigma^2_{\hat{\lambda}}\), but the results we present used
\(\sigma^2=1\). \hfill\break

We also evaluated the precision of \(\hat{\lambda}\) when estimated in
PGLS regression and ANOVA (i.e., \(Y\sim{X}\)). For these simulations,
\(Y=\beta{X}+\epsilon\) was used to generate data with varied
phylogenetic signal and variable associations (\emph{sensu} Revell
2010). In the case of PGLS regression, \(X\) was a variable simulated
under a BM model of evolution, constrained by rescaling phylogeny
according to \(\lambda\), and in the case of ANOVA, \(X\) was a balanced
binary variable (values of 0 and 1), randomly ordered. The strength of
variable associations ranged as \(\beta=0.0,0.25, 0.5, 0.75\) and
\(1.0\). Residuals, \(\epsilon\), were simulated under a BM model of
evolution, constrained by rescaling phylogeny according to \(\lambda\).
Simulated independent variables or residuals were generated from a
normal distribution, i.e.,
\(\mathcal{N}(\mu=0,\sigma=\sigma^2\mathbf{C})\), with the phylogenetic
covariance matrix, \(\mathbf{C}\), implicitly varied for each tree by
the level of \(\lambda\). The fit of the phylogenetic regression was
estimated using maximum likelihood, and parameter estimates
(\(\hat{\lambda}\) and \(\hat{\beta}\)) were obtained. We then
calculated precision estimates (\(\sigma^2_{\hat{\lambda}}\)) at each
input level of phylogenetic signal and level of species richness. We
verified that the level of \(\sigma^2\) had no effect on
\(\sigma^2_{\hat{\lambda}}\) but did influence the precision of
\(\hat{\beta}\) estimated from the linear model (precision increased
with smaller \(\epsilon\), as expected). The results we present used
\(\sigma^2=1\). \hfill\break

All analyses were performed in R v3.6.0 (R Core Team 2019) using the
packages \texttt{geiger} (Harmon et al. 2008), \texttt{caper} (Orme et
al. 2013), \texttt{phytools} (Revell 2012), and \texttt{geomorph} 3.2.1
(Adams and Otárola-Castillo 2013; Adams et al. 2020). R-scripts are
found in the Supporting Information. \hfill\break

\emph{Results.} We found that the precision of \(\lambda_{est}\) varied
widely across simulation conditions. Predictably, precision improved as
the number of species increased (Figure 2). This confirmed earlier
findings of Boettiger et al. (2012), and adhered to parametric
statistical theory. However, in many cases the set of \(\lambda_{est}\)
spanned nearly the entire range of possible values (e.g., \(n=32\);
\(\lambda_{in}=0.5\): \(\lambda_{est}= 0.0\to 0.985\)), revealing that
estimates of \(\lambda\) were not a reliable indicator of input
phylogenetic signal. Importantly, the precision of \(\lambda_{est}\) was
not uniform across all levels of phylogenetic signal. The worst
precision was observed at intermediate levels of phylogenetic signal
(\(\lambda_{in}\approx0.5\)), while precision improved as input levels
approached the extremes of \(\lambda\)'s range (i.e.,
\(\lambda_{in}\rightarrow0\) \& \(\lambda_{in}\rightarrow1\)). Thus,
estimates of \(\lambda\) were least reflective of the true input signal
at intermediate values. Additionally, even at large levels of species
richness, we found that the range of \(\lambda_{est}\) still encompassed
a substantial portion of possible values (e.g., \(n=512\);
\(\lambda_{in}=0.5\): \(\lambda_{est} = 0.32\to 0.68\)). Likewise, the
same \(\lambda_{est}\) could be obtained from datasets containing vastly
different input levels of phylogenetic signal (e.g., \(n=512\);
\(\lambda_{est} = 0.5\); \(\lambda_{in} = 0.25 \to 0.65\)). These
findings were particularly unsettling when considered in light of our
literature survey. Over one quarter of the \(\lambda\) estimates
published in empirical studies (421 of 1,572) were between
\(\lambda=0.25\) and \(\lambda=0.75\) (Figure 1). This range reflected
the region that our simulations identified as being the least reliable
in terms of accurately characterizing levels of phylogenetic signal, yet
30\% of these mid-range empirical estimates were explicitly interpreted
in terms of the strength of phylogenetic signal that they represented
(i.e., weak, intermediate, strong phylogenetic signal). \hfill\break

When phylogenetic signal was investigated on balanced or pectinate
trees, patterns in the precision of \(\lambda\) were largely the same,
with decreased precision at intermediate levels of phylogenetic signal
(Supporting Information). Likewise, when \(\lambda\) was co-estimated
with regression parameters in PGLS regression and ANOVA, the results of
our simulations were quite similar. Regression parameters (\(\beta\))
were accurately estimated, confirming earlier findings of Revell (2010)
(see Supporting Information). However, estimates of phylogenetic signal
(\(\lambda\)) were less precise (Figure 3; see also Supporting
Information), and the spread of \(\lambda_{est}\) was similar to that
observed when \(\lambda\) was estimated for only the dependent variable,
as in Figure 2. Taken together, these findings reveal that
\(\lambda_{est}\) does not precisely characterize known levels of
phylogenetic signal in phenotypic datasets, and that biological
interpretations of the strength of phylogenetic signal based on
\(\lambda\) may be highly inaccurate. \hfill\break

{[}insert Figure 2 here{]} \hfill\break

{[}insert Figure 3 here{]} \hfill\break

\hypertarget{a-standardized-effect-size-for-phylogenetic-signal}{%
\subsection{\texorpdfstring{\emph{A Standardized Effect Size for
Phylogenetic
Signal}}{A Standardized Effect Size for Phylogenetic Signal}}\label{a-standardized-effect-size-for-phylogenetic-signal}}

The results above demonstate that \(\lambda\) is not a reliable estimate
of the phylogenetic signal in phenotypic data. As such, biological
interpretations of the strength of phylogenetic signal, and comparisons
of the magnitude of such effects across datasets are severely
compromised when based on this parameter. As an alternative, we propose
that summary estimates of phylogenetic signal be converted to effect
sizes for interpretation and comparison of the relative strength of
phylogenetic signal in phenotypic datasets. Statistically, a
standardized effect size may be found as:

\begin{align}
    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
\end{align}

where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
its expected value under the null hypothesis, and \(\sigma_\theta\) is
its standard error (Glass 1976; Cohen 1988; Rosenthal 1994).
\(Z_{\theta}\) expresses the magnitude of the effect in \(\theta_{obs}\)
by transforming the original test statistic to its standard normal
deviate (Glass 1976; Kelley and Preacher 2012). Typically,
\(\theta_{obs}\) and \(\sigma_\theta\) are estimated from the data,
while \(E(\theta)\) is obtained from the distribution of \(\theta\)
derived from parametric theory. However, recent advances in resampling
theory (Collyer et al. 2015; Adams and Collyer 2016, 2019a; Collyer and
Adams 2018) have shown that \(E(\theta)\) and \(\sigma_\theta\) may also
be obtained from an empirical sampling distribution of \(\theta\)
obtained from permutation procedures. \hfill\break

Adams and Collyer (2019b) suggested that the strength of phylogenetic
signal could be represented as an effect size based on the \(\kappa\)
statistic and its empirical sampling distribution from permutation. Here
we formalize that suggestion, resulting in an effect size of:

\begin{align}
    Z_\kappa=\frac{log(\kappa_{obs})-\hat\mu_{log(\kappa)}}{\hat\sigma_{log(\kappa)}},
\end{align}

where \(\kappa_{obs}\) is the observed phylogenetic signal, and
\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
deviation of the empirical sampling distribution of \(log(\kappa)\)
obtained via permutation. Note that the logarithm was used because
\(\kappa\) takes only positive values (\(0\to\infty\)) and its sampling
distribution is log-normally distributed (for a similar transformation
when calculating multivariate effect sizes see: Appendix 1 of Collyer
and Adams 2018). \hfill\break

An effect size based on \(\lambda\) is a little more challenging
because, as a statistic, it does not have a sampling distribution on
which to measure standard error. Confidence intervals for \(\lambda\)
are not generated from standard error calculations, but rather for the
values of lambda that intersect the log-likehihood profile for
corresponding percentiles of the \(\chi^2\) distribution used to compare
the compare the putative appropriate model to a null model with
\(\lambda = 0\) {[}citation needed, think Boettiger has a paper on
this{]}. The simplest similar effect size is

\begin{align}
   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
\end{align}

where, \(\hat{\lambda}\) is the maximized likelihood value of
\(\lambda\) and \(\chi^{2}_{\hat{\lambda}}\) is the likelihood ratio
statistic for the value. Hypothetically, this equation would be better
represented as \(Z_{\lambda} = d \sqrt{\chi^2_{\hat{\lambda}}}\), where
\(d\) is a binary value \((-1,1)\) to indicate a direction based on
whether \(\hat{\lambda}\) is below or above a critical value of
\(\lambda\), for a quantile from a \(\chi^{2}\) distribution at a
probability of 0.5. The critical value would correspond to a value of
\(Z = 0\); i.e., the strength of phylogenetic signal that might be
observed at the 50th percentile in a distribution of random outcomes,
generated from a model where \(\lambda = 0\). However, in some
preliminary simulations, we found that mapping
\(\chi^2_{\hat{\lambda}}\) values this way did not produce effect sizes
that were symmetrical about \(Z = 0\), as the mapping is not linear and,
especially for small trees, the log-likelihood profiles can be rather
flat, making estimation of \(\hat{\lambda}\) rather imprecise (as we
elaborate more below.) We will heneforth use equation 3, recognizing
that large values of \(Z\) would likely be positive, anyway, based on
the reference model for the likelihood ratio test.

In this case, \(\lambda_{obs}\) and \(\hat\sigma_\lambda\) are
empirically derived using maximum likelihood, as permutation approaches
have not been developed for evaluating \(\lambda\). Note also that under
the null hypothesis, no phylogenetic signal is expected (Freckleton et
al. 2002), and thus \(E(\lambda)=0\) under this condition. \hfill\break

To evaluate the utility of \(Z_\kappa\) and \(Z_\lambda\) we calculated
both effect sizes for the simulated datasets generated above, and
summarized the precision of each using its variance
(\(\sigma^2_{Z_\kappa}\) and \(\sigma^2_{Z_\lambda}\), Figure 4:
additional results in the Supporting Information). Here two things are
evident. First, estimates of \(Z_\kappa\) linearly track the input
phylogenetic signal whereas estimates of \(Z_\lambda\) do not (Figure
4A, B). Thus, actual changes in the strength of phylogenetic signal are
reflected more evenly in the corresponding values of the effect size
\(Z_\kappa\). Second, the precision of \(Z_\kappa\) is considerably more
stable as compared with \(Z_\lambda\). This may be seen by calculating
the coefficients of variation for the set of precision estimates (i.e.,
\(\sigma^2_{Z_\kappa}\) and \(\sigma^2_{Z_\lambda}\)) across input
levels of phylogenetic signal. Coefficients of variation in the
precision of \(Z_\kappa\) were up to an order of magnitude smaller than
for \(Z_\lambda\) (Figure 4C), implying that estimates of the strength
of phylogenetic signal were more reliable and robust when using
\(Z_\kappa\). \hfill\break

{[}insert Figure 4 here{]}

\hypertarget{statistical-comparisons-of-phylogenetic-signal}{%
\subsection{\texorpdfstring{\emph{Statistical Comparisons of
Phylogenetic
Signal}}{Statistical Comparisons of Phylogenetic Signal}}\label{statistical-comparisons-of-phylogenetic-signal}}

Once the magnitude of phylogenetic signal is characterized using
\(Z_\kappa\), one may wish to compare such measures across datasets, to
determine whether the strength of phylogenetic signal is significantly
greater in one phenotypic trait than another. As with other effect sizes
derived from permutation distributions (e.g., Adams and Collyer 2016,
2019a), a two-sample test statistic may be calculated as:

\begin{align}
  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
\end{align}

where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
side of the equation illustrates that if \(Z_\kappa\) has already been
calculated for two sampling distributions as in equation 2, the sampling
distributions have unit variance for each of the \(Z_\kappa\)
statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
obtained from a standard normal distribution. Typically,
\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
(one-tailed) tests may be specified should the empirical situation
require it (see Adams and Collyer 2016, 2019a). \hfill\break

\hypertarget{empirical-example}{%
\subsection{\texorpdfstring{\emph{Empirical
Example}}{Empirical Example}}\label{empirical-example}}

To demonstrate the utility of \(\hat{Z}_{12}\) we quantified and
compared the strength of phylogenetic signal of two phenotypic traits
across species of plethodontid salamander. The data were part of a
series of studies examining macroevolutionary trends in phenotypic
diversification in this group (Baken and Adams 2019; Baken et al. 2020).
Our dataset contained surface area to volume ratios (SA:V) and relative
body width (\(\frac{BW}{SVL}\)) for 305 species (Figure 5A). For SA:V,
11 linear body measurements from 2,781 individuals were taken, from
which estimates of the surface area and volume of the head, body, and
tail were calculated and subsequently combined to arrive at the SA:V for
each indiviual (for mathematical details see Baken et al. 2020). Species
means were then obtained. Likewise, body size (SVL) and body width (BW)
measurements were taken from 3,371 individuals, and species means of
relative body width (\(\frac{BW}{SVL}\)) were calculated (data from
Baken and Adams 2019). A time-dated molecular phylogeny for the group
(Bonett and Blair 2017) was then pruned to match the species in the
phenotypic dataset. The phylogenetic signal in each trait was then
characterized using \(\kappa\), which was converted to its effect size
(\(Z_\kappa\)) using \texttt{geomorph} 3.2.1 (Adams and Otárola-Castillo
2013; Adams et al. 2020). Finally, the strength of phylogenetic signal
was compared across traits using \(\hat{Z}_{12}\) as described above
(\textbf{to be incorporated in \texttt{geomorph} upon manuscript
acceptance}). \hfill\break

\emph{Results.} Both SA:V and relative body width displayed significant
phylogenetic signal (\(\kappa_{SA:V}=0.7608\); \(P=0.001\);
\(\kappa_{BW/SVL}=0.2515\); \(P=0.001\)). For both phenotypic traits,
\(\kappa_{obs}\) differed markedly from their corresponding permutation
distributions, which were found to overlap almost perfectly (Figure 5B).
However, while both traits displayed significant phylogenetic signal,
there was nearly a four-fold difference in the magnitude of their effect
sizes, with SA:V displaying the greater phylogenetic signal (Figure 5C).
Using the two-sample test statistic above, this difference was found to
be highly significant (\(\hat{Z}_{12}=4.13\); \(P=0.000036\)). Thus it
may be concluded that SA:V displays significantly stronger phylogenetic
signal than does relative body width, and that shared evolutionary
history has strongly influenced trait covariation among taxa for SA:V.
Biologically, this observation corresponds with the fact that tropical
species -- which form a monophyletic group within plethodontids --
display greater variation in SA:V which covaries with disparity in their
climatic niches (Baken et al. 2020). We hypothesize that because of this
macroevolutionary association, strong phylogenetic signal in SA:V is
observed.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

It is common in comparative evolutionary studies to characterize the
phylogenetic signal in phenotypic traits to determine the extent to
which shared evolutionary history has generated trait covariation among
taxa. However, while numerous analytical approaches may be used to
quantify phylogenetic signal (e.g., Abouheif 1999; Gittleman and Kot
1990; Pagel 1999; Blomberg et al. 2003; Adams 2014a), methods that
explicitly measure the strength of phylogenetic signal, or facilitate
comparisons among datasets, have remained underdeveloped. In this study,
we evaluated the precision of one common measure, Pagel's \(\lambda\),
and explored its efficacy for characterizing the strength of
phylogenetic signal in phenotypic data. Using computer simulations, we
found that the precision of \(\lambda\) increased with increasing sample
sizes; a pattern noted previously (Boettiger et al. 2012), and one that
conformed with parametric statistical theory (Cohen 1988). However, we
also found that vastly different \(\lambda\) estimates could be obtained
from data containing the same level of phylogenetic signal, and that
similar \(\lambda\) estimates may be obtained from data containing
differing levels of phylogenetic signal. Further, the precision of
\(\lambda\) varied with the strength of phylogenetic signal, where lower
precision was observed when in data whose phylogenetic signal was of
intemediate strength. From these findings we conclude that \(\lambda\)
is not a reliable indicator of the observed strength of phylogenetic
signal in phenoytpic datasets, and that biological interpretations of
the strength of signal based on this parameter may innacurately
characterize such effects. \hfill\break

As an alternative, we described a standardized effect size (\(Z\)) for
assessing the strength of phylogenetic signal. \(Z\) expresses the
magnitude of phylogenetic signal as a standard normal deviate, which is
easily interpretable as the strength of phylogenetic signal relative to
the mean. We applied this concept to both \(\lambda\) and \(\kappa\),
and found that \(Z_\kappa\) was a better estimate of the strength of
phylogenetic signal in phenotypic data. First, \(Z_\kappa\) was more
precise than \(Z_\lambda\), and precision was more consistent across the
range of input levels of phylogenetic signal. Additionally, values of
\(Z_\kappa\) more accurately tracked known changes in the magnitude of
phylogenetic signal, as demonstrated by the linear relationship between
\(Z_\kappa\) and \(\lambda_{in}\). Thus, \(Z_\kappa\) holds promise as a
measure of the relative strength of phylogenetic signal that reflects
the magnitude of this effect in phenotypic data. We therefore recommend
that future studies interested in the strength of phylogenetic signal
incorporate \(Z_\kappa\) as a statistical measure of this effect.
\hfill\break

Based on the effect size \(Z_\kappa\), we then proposed a two-sample
test, which provides means of determining whether the strength of
phylogenetic signal is greater in one phenotypic trait as compared to
another, via a hypothesis test. Prior studies have summarized patterns
of variation in phylogenetic signal across datasets using summary test
values, such as \(\kappa\) (e.g., Blomberg et al. 2003). However,
\(\kappa\) does not scale linearly with input levels of phylogenetic
signal, and its variance increases (i.e., precision decreases) with
increasing strength of phylogenetic signal (Münkemüller et al. 2012;
Diniz-Filho et al. 2012: see also Supporting Information). Thus,
\(\kappa\) should not be considered an effect size that measures the
strength of phylogenetic signal on a common scale. By contrast,
standardizing \(\kappa\) (\(Z_\kappa\), via equation 2) alleviates these
concerns, and facilitates formal statistical comparisons of the strength
of signal across datasets. Thus when viewed from this perspective, the
approach developed here aligns well with other statistical approaches
such as meta-analysis (sensu Hedges and Olkin 1985; Glass 1976; Arnqvist
and Wooster 1995), where summary statistics across datasets are
converted to standardized effect sizes for subsequent ``higher order''
statistical summaries or comparisons. As such, our approach enables
evolutionary biologists to quantitatively examine the relative strength
of phylogenetic signal across a wide range of phenotypic traits, and
thus opens the door for future discoveries that inform on how phenotypic
diversity accumulates in macroevolutionary time across the tree of life.
\hfill\break

One important advantage of the approach advocated here is that the
resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
measurement cancel out during the calculation of \(Z\) (Sokal and Rohlf
2012). Thus, \(Z_\kappa\) represents the strength of phylogenetic signal
on a common and comparable scale -- measured in standard deviations --
regardless of the initial units and original scale of the phenotypic
variables under investigation. This means that the strength of
phylogenetic signal may be compared across datsets for continuous
phenotypic traits measured in different units and scale, because those
units have been standardized through their conversion to \(Z_\kappa\).
For example, our approach could be utilized to determine whether the
strength of phylogenetic signal (say, in response to ecological
differentiation) is stronger in morphological traits (linear traits:
\(mm\)), physiological traits (metabolic rate: \(\frac{O^2}{min}\)), or
behavioral traits (aggression: \(\frac{\#{displays}}{second}\)). In
fact, our empirical example provided such a comparison, as SA:V is
represented in \(mm^{-1}\) while relative body size is a unitless ratio
(\(\frac{BW}{SVL}\)). Additionally, our method is capable of comparing
the strength of phylogenetic signal in traits of different
dimensionality, as estimates of phylogenetic signal using \(\kappa\)
have been generalized for multivariate data (\(\kappa_{mult}\): see
Adams 2014a). Furthermore, tests based on \(\hat{Z}_{12}\) may be
utilized for comparing the strength of phylogenetic signal among
datasets containing a different number of species, and even for
phenotypes obtained from species in different lineages, because their
phylogenetic non-independence and observed variation are taken into
account in the generation of the empirical sampling distribution via
permutation. \hfill\break

This study is not the first to compare \(\lambda\) and \(\kappa\) for
their ability as statistics to measure phylogenetic signal. Our results
for \(\lambda\) and \(\kappa\) values are consistent with those found in
the simulations performed by Münkemüller et al. (2012), but that study
investigated type I error rates and statistical power, finding that
\(\lambda\) performed better in both regards, irrespective of species
number in trees. Although not the central focus of their study, the same
tendency for variable \(\lambda\) and consistent \(\kappa\) at
intermediate phylogenetic signal strengths was observed (see Fig. 2,
Münkemüller et al. 2012). Recent work by Molina-Venegas and Rodríguez
(2017) found that \(\kappa\) but not \(\lambda\) tended to inflate the
estimate of phylogenetic signal, leading to moderate type I and type II
biases, if polytomic chronograms were used. Their work more thoroughly
addressed previous observations of inflated \(\kappa\) for incompletely
resolved phylogenetic trees (Davies et al. 2012; Münkemüller et al.
2012). An interesting question is whether an inflated \(\kappa\) value
leads to an inflated \(Z_\kappa\) or does a tendency of a particular
tree to inflate estimates of \(\kappa\) also inflate the values in
random permutations of a test, in which case \(Z_\kappa\) is robust to
polytomies? We repeated the analyses in Figure 4, adjusting trees to
have 50\% collapsed nodes, per the technique of Molina-Venegas and
Rodríguez (2017), and found results were consistent (see Supporting
Information). This confirms that any tendency of incompletely resolved
trees to inflate \(\kappa\) as a descriptive statistic does not inflate
\(Z_\kappa\) as an effect size. Furthermore, because comparison of
effect sizes in a test is a comparison of locations of observed values
in their sampling distributions, which would shift concomitantly because
of this tendency, the \(Z_{12}\) test statistic in equation 4 appears to
be robust in spite of unresolved trees. \hfill\break

Phylogenetic signal can be thought of as both an attribute to be
measured in the data and a parameter that can be tuned to account for
the phylogenetic non-independence among observations, for analysis of
the data. As such, \(\lambda\) is appealing, as a statistic that
potentially fulfills both roles. However, the inability to estimate
phylogenetic signal with \(\lambda\) for data simulated with known
phylogenetic signal is troublesome, and we recommend evolutionary
biologists refrain from viewing it as a statistic to describe the amount
of phylogenetic signal in the data. Interestingly, \(\kappa\) -- when
standardized to an effect size \(Z_\kappa\) -- is a better statistic for
measuring the amount of phylogenetic signal in data simulated with
respect to known levels of \(\lambda\). Although \(\lambda\) might be
viewed as an important parameter for modifying the the conditional
estimation of linear model coefficients with respect to phylogeny, it is
neither a statistic that has meaningful comparative value as a measure
of phylogenetic signal nor a statistic that lends itself well to
reliable calculation of a test statistic. By contrast, \(\kappa\) has
been shown here to be a reliable statistic, but only when standardized
by the mean and standard deviation of its empirical sampling
distribution (i.e., when converted to the effect size, \(Z_\kappa\)).
Because one has control over the number of permutations used in
analysis, one can be assured with many permutations that the empirical
sampling distribution is representative of true probability
distributions (Adams and Collyer 2018). With low coefficients of
variation for \(Z_\kappa\) (Figure 4), it is difficult to imagine that a
hypothesis test can improve equation 4 for efficiently comparing
phylogenetic signal for different traits, different trees, or a
combination of both.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\setlength{\parindent}{-0.25in} \setlength{\leftskip}{0.25in}
\setlength{\parskip}{8pt} \noindent

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abouheif1999}{}%
Abouheif, E. 1999. A method for testing the assumption of phylogenetic
independence in comparative data. Evolutionary Ecology Research
1:895--909.

\leavevmode\hypertarget{ref-Adams2014a}{}%
Adams, D. C. 2014a. A generalized Kappa statistic for estimating
phylogenetic signal from shape and other high-dimensional dultivariate
data. Systematic Biology 63:685--697.

\leavevmode\hypertarget{ref-Adams2014b}{}%
Adams, D. C. 2014b. A method for assessing phylogenetic least squares
models for shape and other high-dimensional multivariate data. Evolution
68:2675--2688.

\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
Adams, D. C., and M. L. Collyer. 2019a. Comparing the strength of
modular signal, and evaluating alternative modular hypotheses, using
covariance ratio effect sizes with morphometric data. Evolution
73:2352--2367.

\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
Adams, D. C., and M. L. Collyer. 2016. On the comparison of the strength
of morphological integration across morphometric datasets. Evolution
70:2623--2631.

\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
Adams, D. C., and M. L. Collyer. 2018. Phylogenetic anova: Group-clade
aggregation, biological challenges, and a refined permutation procedure.
Evolution 72:1204--1215.

\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
Adams, D. C., and M. L. Collyer. 2019b. Phylogenetic comparative methods
and the evolution of multivariate phenotypes. Annual Review of Ecology,
Evolution, and Systematics 50:405--425.

\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
Adams, D. C., M. L. Collyer, and A. Kaliontzopoulou. 2020. Geomorph:
Software for geometric morphometric analyses. R package version 3.2.1.

\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
Adams, D. C., and E. Otárola-Castillo. 2013. Geomorph: An r package for
the collection and analysis of geometric morphometric shape data.
Methods in Ecology and Evolution 4:393--399.

\leavevmode\hypertarget{ref-Arnqvist1995}{}%
Arnqvist, G., and D. Wooster. 1995. Meta-analysis: Synthesizing research
findings in ecology and evolution. Trends in Ecology and Evolution
10:236--240.

\leavevmode\hypertarget{ref-Bai2019}{}%
Bai, K., S. Lv, S. Ning, D. Zeng, Y. Guo, and B. Wang. 2019. Leaf
nutrient concentrations associated with phylogeny, leaf habit and soil
chemistry in tropical karst seasonal rainforest tree species. Plant and
Soil 434:305--326.

\leavevmode\hypertarget{ref-Baken2019}{}%
Baken, E. K., and D. C. Adams. 2019. Macroevolution of arboreality in
salamanders. Ecology and Evolution 9:7005--7016.

\leavevmode\hypertarget{ref-Baken2020}{}%
Baken, E. K., L. E. Mellenthin, and D. C. Adams. 2020. Macroevolution of
desiccation‐related morphology in plethodontid salamanders as inferred
from a novel surface area to volume ratio estimation approach. Evolution
74:476--486.

\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
Beaulieu, J. M., D. C. Jhwueng, C. Boettiger, and B. C. O'Meara. 2012.
Modeling stabilizing selection: Expanding the ornstein-uhlenbeck model
of adaptive evolution. Evolution 66:2369--2383.

\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
Blomberg, S. P., T. Garland, and A. R. Ives. 2003. Testing for
phylogenetic signal in comparative data: Behavioral traits are more
labile. Evolution 57:717--745.

\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
Boettiger, C., G. Coop, and P. Ralph. 2012. Is your phylogeny
informative? Measuring the power of comparative methods. Evolution
67:2240--2251.

\leavevmode\hypertarget{ref-Bonett2017}{}%
Bonett, R. M., and A. L. Blair. 2017. Evidence for complex life cycle
constraints on salamander body form diversification. Proceedings of the
National Academy of Sciences, U.S.A. 114:9936--9941.

\leavevmode\hypertarget{ref-Bose2019}{}%
Bose, R., B. R. Ramesh, R. Pélissier, and F. Munoz. 2019. Phylogenetic
diversity in the western ghats biodiversity hotspot reflects
environmental filtering and past niche diversification of trees. Journal
of Biogeography 46:145--157.

\leavevmode\hypertarget{ref-ButlerKing2004}{}%
Butler, M. A., and A. A. King. 2004. Phylogenetic comparative analysis:
A modeling approach for adaptive evolution. American Naturalist
164:683--695.

\leavevmode\hypertarget{ref-Cohen1988}{}%
Cohen, J. 1988. Statistical power analysis for the behavioral sciences.
Routledge.

\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
Collyer, M. L., and D. C. Adams. 2018. RRPP: An r package for fitting
linear models to high-dimensional data using residual randomization.
Methods in Ecology and Evolution 9:1772--1779.

\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
Collyer, M. L., D. J. Sekora, and D. C. Adams. 2015. A method for
analysis of phenotypic change for phenotypes described by
high-dimensional data. Heredity 115:357--365.

\leavevmode\hypertarget{ref-Cooper2010}{}%
Cooper, N., W. Jetz, and R. P. Freckleton. 2010. Phylogenetic
comparative approaches for studying niche conservatism. Journal of
Evolutionary Biology 23:2529--2539.

\leavevmode\hypertarget{ref-Davies2012}{}%
Davies, T. J., N. J. Kraft, N. Salamin, and E. M. Wolkovich. 2012.
Incompletely resolved phylogenetic trees inflate estimates of
phylogenetic conservatism. Ecology 93:242--247. Wiley Online Library.

\leavevmode\hypertarget{ref-DeMeester2019}{}%
De Meester, G., K. Huyghe, and R. Van Damme. 2019. Brain size, ecology
and sociality: A reptilian perspective. Biological Journal of the
Linnean Society 126:381--391.

\leavevmode\hypertarget{ref-DinizFilho2012}{}%
Diniz-Filho, J. A. F., T. Santos, T. F. Rangel, and L. M. Bini. 2012. A
comparison of metrics for estimating phylogenetic signal under
alternative evolutionary models. Genetics and Molecular Biology
35:673--679.

\leavevmode\hypertarget{ref-Felsenstein1985}{}%
Felsenstein, J. 1985. Phylogenies and the comparative method. American
Naturalist 125:1--15.

\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
Freckleton, R. P., P. H. Harvey, and M. Pagel. 2002. Phylogenetic
analysis and comparative data: A test and review of evidence. American
Naturalist 160:712--726.

\leavevmode\hypertarget{ref-GarlandIves2000}{}%
Garland, T. J., and A. R. Ives. 2000. Using the past to predict the
present: Confidence intervals for regression equations in phylogenetic
comparative methods. American Naturalist 155:346--364.

\leavevmode\hypertarget{ref-Gittleman1990}{}%
Gittleman, J. L., and M. Kot. 1990. Adaptation: Statistics and a null
model for estimating phylogenetic effects. Systematic Zoology
39:227--241.

\leavevmode\hypertarget{ref-Glass1976}{}%
Glass, G. V. 1976. Primary, secondary, and meta-analysis of research.
Educational Researcher 5:3--8.

\leavevmode\hypertarget{ref-Grafen1989}{}%
Grafen, A. 1989. The phylogenetic regression. Philosophical Transactions
of the Royal Society of London B, Biological Sciences 326:119--157.

\leavevmode\hypertarget{ref-Harmon2008}{}%
Harmon, L. J., J. T. Weir, C. D. Brock, R. E. Glor, and W. Challenger.
2008. GEIGER: Investigating evolutionary radiations. Bioinformatics
24:129--131.

\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
Harvey, P. H., and M. D. Pagel. 1991. The comparative method in
evolutionary biology. Oxford University Press, Oxford.

\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
Hedges, L. V., and I. Olkin. 1985. Statistical methods for
meta-analysis. Elsevier.

\leavevmode\hypertarget{ref-Kelley2012}{}%
Kelley, K., and K. J. Preacher. 2012. On effect size. Psychological
Methods 17:137--152.

\leavevmode\hypertarget{ref-Klingenberg2010}{}%
Klingenberg, C. P., and N. A. Gidaszewski. 2010. Testing and quantifying
phylogenetic signals and homoplasy in morphometric data. Systematic
biology 59:245--261.

\leavevmode\hypertarget{ref-Liu2019}{}%
Liu, H., C. P. Osborne, D. Yin, R. P. Freckleton, G. Jiang, and M. Liu.
2019. Phylogeny and ecological processes influence grass coexistence at
different spatial scales within the steppe biome. Oecologia 191:25--38.

\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
Martins, E. P., and T. F. Hansen. 1997. Phylogenies and the comparative
method: A general approach to incorporating phylogenetic information
into the analysis of interspecific data. American Naturalist
149:646--667.

\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
Molina-Venegas, R., and M. A. Rodríguez. 2017. Revisiting phylogenetic
signal; strong or negligible impacts of polytomies and branch length
information? BMC evolutionary biology 17:53.

\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
Münkemüller, T., S. Lavergne, B. Bzeznik, S. Dray, T. Jombart, K.
Schiffers, and W. Thuiller. 2012. How to measure and test phylogenetic
signal. Methods in Ecology and Evolution 3:743--756.

\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
O'Meara, B. C., C. Ane, M. J. Sanderson, and P. C. Wainwright. 2006.
Testing for different rates of continuous trait evolution using
likelihood. Evolution 60:922--933.

\leavevmode\hypertarget{ref-Orme2013}{}%
Orme, D., R. Freckleton, G. Thomas, T. Petzoldt, S. Fritz, and N. Isaac.
2013. CAPER: Comparative analyses of phylogenetics and evolution in r.
Methods in Ecology and Evolution 3:145--151.

\leavevmode\hypertarget{ref-Pagel1999}{}%
Pagel, M. D. 1999. Inferring the historical patterns of biological
evolution. Nature 401:877--884.

\leavevmode\hypertarget{ref-Pavoine2012}{}%
Pavoine, S., and C. Ricotta. 2012. Testing for phylogenetic signal in
biological traits: The ubiquity of cross-product statistics. Evolution:
International Journal of Organic Evolution 67:828--840.

\leavevmode\hypertarget{ref-Pintanel2019}{}%
Pintanel, P., M. Tejedo, S. R. Ron, G. A. Llorente, and A.
Merino-Viteri. 2019. Elevational and microclimatic drivers of thermal
tolerance in andean pristimantis frogs. Journal of Biogeography
46:1664--1675.

\leavevmode\hypertarget{ref-RCT}{}%
R Core Team. 2019. R: A language and environment for statistical
computing. R Foundation for Statistical Computing, Vienna, Austria.

\leavevmode\hypertarget{ref-Revell2010}{}%
Revell, L. J. 2010. Phylogenetic signal and linear regression on species
data. Methods in Ecology and Evolution 1:319--329.

\leavevmode\hypertarget{ref-Revell2012}{}%
Revell, L. J. 2012. Phytools: An r package for phylogenetic comparative
biology (and other things). Methods in Ecology and Evolution 3:217--223.

\leavevmode\hypertarget{ref-RevellHarmon2008}{}%
Revell, L. J., and L. J. Harmon. 2008. Testing quantitative genetic
hypotheses about the evolutionary rate matrix for continuous characters.
Evolutionary Ecology Research 10:311--331.

\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
Revell, L. J., L. J. Harmon, and D. C. Collar. 2008. Phylogenetic
signal, evolutionary process, and rate. Systematic Biology 57:591--601.

\leavevmode\hypertarget{ref-Rohlf2001}{}%
Rohlf, F. J. 2001. Comparative methods for the analysis of continuous
variables: Geometric interpretations. Evolution 55:2143--2160.

\leavevmode\hypertarget{ref-Rosenthal1994}{}%
Rosenthal, R. 1994. The handbook of research synthesis. Pp. 231--244
\emph{in} L. V. Cooper H Hedges, ed. Russell Sage Foundation.

\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
Sokal, R. R., and F. J. Rohlf. 2012. Biometry. 4th ed. W.H. Freeman \&
Co., San Francisco.

\leavevmode\hypertarget{ref-Su2019}{}%
Su, G., S. Villéger, and S. Brosse. 2019. Morphological diversity of
freshwater fishes differs between realms, but morphologically extreme
species are widespread. Global ecology and biogeography 28:211--221.

\leavevmode\hypertarget{ref-Vandelook2019}{}%
Vandelook, F., S. Janssens, P. Gijbels, E. Fischer, W. Van den Ende, O.
Honnay, and S. Abrahamczyk. 2019. Nectar traits differ between
pollination syndromes in balsaminaceae. Annals of Botany 124:269--279.

\newpage

\hypertarget{figure-legends}{%
\section{Figure Legends}\label{figure-legends}}

\textbf{Figure 1}. Frequency distribution of \(\lambda\) estimates
published in 2019. The majority of these values were close to 0 or 1,
and from phylogenies with fewer than 200 taxa. \hfill\break

\textbf{Figure 2}. Precision of Pagel's \(\lambda\) across known levels
of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
various sizes. As phylogenies increase in size, variation in
\(\lambda_{in}\) decreases; however the precision is not constant across
the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
intermediate levels of phylogenetic signal. \hfill\break

\textbf{Figure 3}. Precision of Pagel's \(\lambda\) when incorporated in
phylogenetic regression (\(Y\sim X\)), across known levels of input
phylogenetic signal (\(\lambda_{in}\)) on phylogenies of various sizes.
As phylogenies increase in size, variation in \(\lambda_{in}\)
decreases; however the precision is not constant across the range of
input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at intermediate
levels of phylogenetic signal. \hfill\break

\textbf{Figure 4}. Variation in effect size estimates of phylogenetic
signal across input levels of phylogenetic signal. (A) Estimates
\(Z_\lambda\) for data simulated on phylogenies with 32 taxa (\(n=32\)),
(B) Estimates of \(Z_\kappa\) for data simulated on phylogenies with 32
taxa (\(n=32\)), (C) Coefficients of variation of precision estimates of
\(Z_\lambda\) and \(Z_\kappa\) across input levels of phylogenetic
signal, estimated on phylogenies containing differing numbers of
species.

\textbf{Figure 5}. (A) Linear measures for relative body size, and
regions of the body used to estimate surface area to volume (SA:V)
ratios. (B) Permutation distributions of phylogenetic signal for SA:V
and \(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\), with their
95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).

\newpage

\includegraphics[width=0.95\linewidth]{Fig1}

\singlespacing \textbf{Figure 1}. Frequency distribution of \(\lambda\)
estimates published in 2019. The majority of these values were close to
0 or 1, and from phylogenies with fewer than 200 taxa.

\newpage

\includegraphics[width=0.95\linewidth]{Fig2}

\singlespacing \textbf{Figure 2}. Precision of Pagel's \(\lambda\)
across known levels of input phylogenetic signal (\(\lambda_{in}\)) on
phylogenies of various sizes. As phylogenies increase in size, variation
in \(\lambda_{in}\) decreases; however the precision is not constant
across the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is
highest at intermediate levels of phylogenetic signal.

\newpage

\includegraphics[width=0.95\linewidth]{Fig3}

\singlespacing \textbf{Figure 3}. Precision of Pagel's \(\lambda\) when
incorporated in phylogenetic regression (\(Y\sim X\)), across known
levels of input phylogenetic signal (\(\lambda_{in}\)) on phylogenies of
various sizes. As phylogenies increase in size, variation in
\(\lambda_{in}\) decreases; however the precision is not constant across
the range of input levels (\(\lambda_{in}: 0 \to 1\)), and is highest at
intermediate levels of phylogenetic signal.

\newpage

\includegraphics[width=0.95\linewidth]{Fig4}

\singlespacing \textbf{Figure 4}. Variation in effect size estimates of
phylogenetic signal across input levels of phylogenetic signal. (A)
Estimates \(Z_\lambda\) for data simulated on phylogenies with 32 taxa
(\(n=32\)), (B) Estimates of \(Z_\kappa\) for data simulated on
phylogenies with 32 taxa (\(n=32\)), (C) Coefficients of variation of
precision estimates of \(Z_\lambda\) and \(Z_\kappa\) across input
levels of phylogenetic signal, estimated on phylogenies containing
differing numbers of species.

\includegraphics[width=0.95\linewidth]{Fig5}

\singlespacing \textbf{Figure 5}. (A) Linear measures for relative body
size, and regions of the body used to estimate surface area to volume
(SA:V) ratios. (B) Permutation distributions of phylogenetic signal for
SA:V and \(\frac{BW}{SVL}\), with observed values shown as vertical
bars. (C) Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\),
with their 95\% confidence intervals (CI not standardized by
\(\sqrt(n)\)).

\end{document}