### 2020-07-16 Manuscript/PNAS-Adams-01.tex

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment.

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\templatetype{pnasresearcharticle}  % Choose template

\title{A Standardized Effect Size for Evaluating and Comparing the Strength of
Phylogenetic Signal}

\author[a,1]{Dean C. Adams}
\author[a,b]{Erica K. Baken}
\author[b]{Michael L. Collyer}

  \affil[a]{Department of Ecology, Evolution, and Organismal Biology, Iowa State
University, Ames, Iowa, 50010. USA.}
  \affil[b]{Department of Science, Chatham University, Pittsburgh, Pennsylvania,
15232. USA.}

% Please give the surname of the lead author for the running footer
\leadauthor{Adams et al.}

% Please add here a significance statement to explain the relevance of your work
\significancestatement{Evolutionary biologists wish to quantify and compare the strength of
phylogenetic signal across datasets, but analytical tools for these
comparisons are generally lacking. Here we develop a standardized effect
size, \(Z_\kappa\), which measures the strength of phylogenetic signal
on a common statistical scale. We also provide a test statistic,
\(\hat{Z}_{12}\), for comparing the strength of phylogenetic signal
across datasets. We find that two commonly used parameters (Pagel's
\(\lambda\) and Blomberg's \(\kappa\)), not converted to effect sizes,
are unsuitable for this purpose. Our effect-size procedure enables
biologists to quantitatively address hypotheses that compare the
strength of phylogenetic signal between various phenotypic traits, even
when those traits are found in different evolutionary lineages or have
different units or scales.}

\authorcontributions{D.C.A. designed the research; D.C.A., E.K.B., and M.L.C. performed the
research and wrote the paper.}

\authordeclaration{The authors declare no conflict of interest. \hfill\break

Data deposition: Data for the empirical example may be found on DRYAD:
\url{doi:10.5061/dryad.b554m44} and \url{doi:10.5061/dryad.59zw3r23m}.
R-scripts for simulation tests are found on Github: XXX. Computer code
for implementing the two-sample comparison of effect sizes is found in
geomorph:
\url{https://cran.r-project.org/web/packages/geomorph/index.html}
\hfill\break

\textsuperscript{1}To whom correspondence should be addressed. E-mail:
\href{mailto:dcadams@iastate.edu}{\nolinkurl{dcadams@iastate.edu}}}

\correspondingauthor{\textsuperscript{} }

% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
 \keywords{  phylogenetic signal |  macroevolution |  lambda |  kappa  }

\begin{abstract}
Macroevolutionary studies frequently characterize the phylogenetic
signal in phenotypes, however, analytical tools for comparing the
strength of that signal across traits remain largely underdeveloped.
Here we evaluate the efficacy of of Pagel's \(\lambda\) to correctly
estimate the strength of phylogenetic signal in phenotypic traits across
a range of input values. We find that \(\lambda\) behaves as a Bernoulli
random variable, where estimates are increasingly skewed at larger and
smaller input levels of phylogenetic signal. Further, the precision of
\(\lambda\) varies with input signal. Another measure, Blomberg's
\(\kappa\), is more consistent across a range of tree sizes, and
exhibits a positive relationship with input levels of phylogenetic
signal. However, that relationship is decidedly nonlinear. Thus, neither
\(\lambda\) nor \(\kappa\) are suitable as effect sizes for measuring
the strength of phylogenetic signal, and comparing that signal across
datasets. As an alternative, we propose a standardized effect size based
on \(\kappa\), (\(Z_\kappa\)), which measures the strength of
phylogenetic signal more reliably than does \(\lambda\), and places that
signal on a common scale for statistical comparison. We develop tests
based on \(Z_\kappa\) to provide a mechanism for formally comparing the
strength of phylogenetic signal across datasets, in much the same manner
as effect sizes may be used to summarize patterns in quantitative
meta-analysis. Our approach extends the phylogenetic comparative toolkit
to address hypotheses that compare the strength of phylogenetic signal
between various phenotypic traits, even when those traits are found in
different evolutionary lineages or have different units or scales.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
% You should only change this length when you've finalised the article contents.
\verticaladjustment{-2pt}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

% If your first paragraph (i.e. with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation. If this is the case, add \parshape=0 to the end of the list environment.

\acknow{We thank E. Glynne and B. Juarez for comments on early drafts of the
manuscript. This work was supported in part by NSF grant DBI-1902511 (to
D.C.A.) and DBI-1902694 (to M.L.C.). \hfill\break}

Investigating macroevolutionary patterns of trait variation requires a
phylogenetic perspective, because the shared ancestry among species
violates the assumption of independence among trait values that is
common for statistical tests (1, 2). Accounting for this evolutionary
non-independence is the purview of \emph{phylogenetic comparative
methods} (PCMs): a suite of analytical tools that condition trends in
the data on the phylogenetic relatedness of observations (3--10). These
methods are predicated on the notion that phylogenetic signal -- the
tendancy for closely related species to display similar trait values --
is present in cross-species datasets (1, 11, 12). Indeed, under numerous
evolutionary models, phylogenetic signal is to be expected, as
stochastic character change along the hierarchical structure of the tree
of life generates trait covaration among related taxa (1, 12, 13).

Several analytical tools have been developed to quantify phylogenetic
signal in phenotypic datasets (11, 12, 14--17), and their statistical
properties -- namely type I error rates and statistical power -- have
been investigated to determine under what conditions phylogenetic signal
can be detected (13, 16, 18--23). One of the most widely used methods
for characterizing phylogenetic signal is Pagel's \(\lambda\) (11),
which transforms the lengths of the internal branches of the phylogeny
to improve the fit of data to the phylogeny via maximum likelihood (11,
24). When incorporated in PGLS, \(\lambda\) serves as a tuning parameter
which is optimized via log-likelihood profiling while evaluating the
covariation between the dependent and independent variables, given the
phylogeny (11, 24). To infer whether phylogenetic signal differs from no
signal or a Brownian motion model of evolutionary divergence, the
observed model fit using \(\hat\lambda\) may be statistically compared
to that using \(\lambda=0\) or \(\lambda=1\) via likelihood ratio tests
(24--26) or confidence limits (27).

Another widely used measure of phylogenetic signal is Blomberg's
\(\kappa\) (12), which characterizes phylogenetic signal as the ratio of
observed trait variation to the amount of variation expected under
Brownian motion. Blomberg's \(\kappa\) can be treated as a test
statistic by employing a permutation test to generate its sampling
distribution (12, 16) for determining whether significant phylogenetic
signal is present in data. Both \(\lambda\) and \(\kappa\) seem
intuitive to interpret, as a value of \(0\) for both corresponds to no
phylogenetic signal, while a value of \(1\) corresponds to the amount of
phylogenetic signal expected under Brownian motion. Thus, it is tempting
to regard both \(\lambda\) and \(\kappa\) as descriptive statistics that
measure the relative strength of phylogenetic signal, providing an
estimate of its magnitude for comparison.

The appeal of Pagel's \(\lambda\) and Blomberg's \(\kappa\) as
descriptive statistics is that they provide a basis for interpreting
``weak'' versus ``strong'' phylogenetic signal; i.e., small versus large
values of \(\hat{\lambda}\) or \(\kappa\), respectively, in a
comparative sense (28--30). Nonetheless, an important question that has
yet to be considered is whether these statistics are or can be converted
to effect sizes for comparative analyses across datasets? To be
statistics representing phylogenetic signal, they should have reliable
distributional properties, which could be revealed with simulation
experiments. For instance, as a proportional random variable bounded by
\(0\) and \(1\), we might expect that \(\hat{\lambda}\) follows a
Bernoulli distribution (\textbf{add ref}); i.e., branch lengths in a
tree are scaled proportionally to the probability that data arise from a
BM process. Given a known \(\lambda\) value used to generate random data
on a tree, we would also expect that the mean of an empirical sampling
distribution of \(\hat{\lambda}\) would approximately equal \(\lambda\);
the dispersion of \(\hat{\lambda}\) would be largest at intermediate
values of \(\lambda\), \(\hat{\lambda}\) would be predictable over the
range of \(\lambda\) with respect to tree size; the distribution of
\(\hat{\lambda}\) would be symmetric at intermediate values of
\(\lambda\) and more skewed toward values of 0 or 1; and that the
distribution of \(\hat{\lambda}\) will be more platykurtic at
intermediate values of \(\lambda\), becoming more leptokurtic toward 0
and 1 (\textbf{add same ref}). Prior work (18) seems to support some of
these conjectures, based superficially on statistical moments for a
given tree size (mean, variance, skewness, and kurtosis; see Fig. 2 of
ref. (18)). However, because the ``strength of Brownian motion'' was
simulated as a varied weighted-average of data simulated on trees with
\(\lambda=0\) and \(\lambda=1\) and not as prescribed values of
\(\lambda\) (18), interpretation of these patterns is challenging.

By contrast, for Blomberg's \(\kappa\), which is positively ubounded, we
might expect that for any \(\lambda\) used to generate data, estimates
of \(\kappa\) might follow a normal distribution, with values
distributed symmetrically about the input value. This attribute seemed
less reasonable based on the simulations performed by Münkemüller et
al.~(18), which suggested that distributions were positively skewed and
that Blomberg's \(\kappa\) might not behave as a statistic that follows
a normal distribution. However, because their simulations used a
weighted combination of simulated phylogenetic signal strengths, strong
inferences are not possible (and distributional attributes were not the
intended result of their simulations). Thus, for both Pagel's
\(\lambda\) or Blomberg's \(\kappa\), evaluation of statistical moments
across a range of \(\lambda\) used to generate data would be valuable
for adjudicating the relibalilty of these statistics. Furthermore, these
are statistics that appear to have expected values that vary with tree
size (18), making comparisons across studies challenging. Therefore,
transformation of these statistics into \(Z\)-scores would allow
evaluation of the efficacy of each statistic to yield effect sizes that
could be used for comparisons of the strength of phylogenetic signal
across traits and lineages.

Here we use simulation experiments to compare the distributional
attributes of \(\hat{\lambda}\) and \(\kappa\), plus their effect sizes
(\(Z\)-scores), across a range of tree size and phylogenetic signal
strength. We find that estimates of \(\hat{\lambda}\) are increasingly
skewed at larger and smaller input levels of phylogenetic signal and at
smaller tree sizes, vary widely for a given input value of \(\lambda\),
and that the precision of \(\hat{\lambda}\) is not constant across its
range. By contrast, estimates of \(\kappa\) are more consistent across
tree sizes, and are normally distributed across the range of input
levels of \(\lambda\), making \(\kappa\) a more reliable statistic. We
then propose an effect size based on \(\kappa\), (\(Z_{\kappa}\)), which
provides consistent estimates of the strength of phylogenetic signal
across tree sizes and signal strength, and facilitates quantitative
comparisons of the relative strength of phylogenetic signal across
datasets.

\hypertarget{results}{%
\section{Results}\label{results}}

\textbf{Lambda (\(\lambda\)) estimates of phylogenetic signal are
innacurate.} Computer simulations reveal that for \(\hat{\lambda}\), the
distributional expectations of a Bernoulli variable were mostly upheld.
First, the mean value of \(\hat{\lambda}\) increases as \(\lambda\)
increases, but it is negatively-biased (particularly for small tree
sizes), and is consistently less than the input \(\lambda\) value across
most of its range (Fig. 1 black line). Second, the standard deviation of
\(\hat{\lambda}\) is largest at intermediate values of \(\lambda\) and
smallest at extreme values, implying that the precision in estimating
\(\lambda\) varies across the range of input values (Fig. 1 red line).
Additionally, standard deviations of \(\hat{\lambda}\) are negatively
associated with tree size, and for tress of 128 species or less,
\(\hat{\lambda}\) are quite variable, except for cases when \(\lambda\)
is near or equal to \(1\). Third, the distributions of \(\hat{\lambda}\)
are not normal across its range, but become increasingly skewed at more
extreme values of \(\lambda\) (Fig. 1 blue line). For small tree sizes,
it is also clear that distributions are more platykurtic at intermediate
values of \(\hat{\lambda}\). Taken together these results reveal that
\(\hat\lambda\) inconsistently estimates phylogenetic signal, both
across tree sizes and across the range of input values. Additional
simulations (Supplemental Information) reveal that incorporating
\(\hat\lambda\) in PGLS anova and regression does not adversely affect
the statistical properties of PGLS parameter estimation or model
evaluation (type I error, power, bias in coefficients). Thus, it is
reasonable to incorporate \(\hat\lambda\) in PGLS as a parameter for
tuning the degree of phylogenetic signal in the dependent variables
during the analysis. However, the statistical properties shown in Fig. 1
demonstrate that \(\lambda\) is unsuitable as an effect size for
measuring the strength of phylogenetic signal in data, or for comparing
that signal across datasets.

\begin{figure}
\centering
\includegraphics{new.fig.1.temp.png}
\caption{Response of Pagel's \(\lambda\) to increasing strength of
Brownian motion. Gray line signifies the 1:1 line where the input value
matches the estimate \(\hat\lambda\). At each input level, the dark
black line represents the empirically derived expected value (mean) of
\(\hat\lambda\), the red line is the standard deviation of
\(\hat\lambda\), and the blue line is Shapiro Wilks statistic of
\(\hat\lambda\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
skewed distributions).{}}
\end{figure}

\textbf{Kappa (\(\kappa\)) estimates of phylogenetic signal are more
reliable.} Simulation results for \(\hat\kappa\) demonstrate that
\(\hat\kappa\) displays better statistical properties. First, as
expected, mean values of \(\hat\kappa\) increase with increasing signal
(\(\lambda\)) irrespective of tree size, though the increase does not
scale linearly with input levels of phylogenetic signal (Fig. 2 black
line). Additionally, the standard deviation of \(\hat\kappa\) is
consistent across tree sizes (Fig. 2 red line), and while it increases
with \(\lambda\), it is always less than the corresponding mean. This
finding is perhaps unsurprising, as \(\hat\kappa\) is lower-bounded by
0, and is never large for small values of \(\lambda\). Importantly,
\(\hat\kappa\) is normally distributed across the range of input
\(\lambda\), and remains consistent in this pattern regardless of tree
size (Fig. 2 blue line). In fact, distributions of \(\hat\kappa\)
display slight skewing only at small tree sizes and when \(\lambda\) is
near or equal to \(1\). This result differs from those of (18), where
skewing appears to be from combining random values generated
independently. These findings reveal that while \(\kappa\) is more
reliable as an estimate of phylogenetic signal, the non-linear scaling
with input signal implies that it should not be considered an effect
size that measures the strength of phylogenetic signal on a common scale
for comparison across datasets.

\begin{figure}
\centering
\includegraphics{new.fig.2.temp.png}
\caption{Response of Blomberg's \(\hat\kappa\) to increasing strength of
Brownian motion. Gray line signifies the 1:1 line where the input value
matches the estimate \(\hat\kappa\). At each input level, the dark black
line represents the empirically derived expected value (mean) of
\(\hat\kappa\), the red line is the standard deviation of
\(\hat\kappa\), and the blue line is Shapiro Wilks statistic of
\(\hat\kappa\) (\(W=1.0\) signifies normality, \(W< 1.0\) represent
skewed distributions).{}}
\end{figure}

\textbf{Effect sizes from \(\kappa\) (\(Z_{\kappa}\)) better
characterize phylogenetic signal.} To measure the strength of
phylogenetic signal on a common scale, here we propose effect sizes
(Z-scores) for both \(\lambda\) and \(\kappa\). Statistically, a
standardized effect size may be found as:

\begin{align}
    Z_{\theta}=\frac{\theta_{obs}-E(\theta)}{\sigma_\theta}
\end{align}

where \(\theta_{obs}\) is the observed test statistic, \(E(\theta)\) is
its expected value under the null hypothesis, and \(\sigma_\theta\) is
its standard error (31--33). Typically, \(\theta_{obs}\) and
\(\sigma_\theta\) are estimated from the data, while \(E(\theta)\) is
obtained from the distribution of \(\theta\) derived from parametric
theory. However, recent advances in resampling theory (34--37) have
shown that \(E(\theta)\) and \(\sigma_\theta\) may also be obtained from
an empirical sampling distribution of \(\theta\) obtained from
permutation procedures.

Formalizing the suggestion of Adams and Collyer (38), an effect size for
\(\kappa\) may be found as:

\begin{align}
    Z_\kappa=\frac{\kappa_{obs}-\hat\mu_{\kappa}}{\hat\sigma_{\kappa}},
\end{align}

where \(\kappa_{obs}\) is the observed phylogenetic signal, and
\(\hat\mu_\kappa\) and \(\hat\sigma_\kappa\) are the mean and standard
deviation of the empirical sampling distribution of \(\kappa\) obtained
via permutation. The empirical sampling distribution of \(\kappa\) is
first transformed via Box-Cox to better adhere to the assumption of
normality.

For \(\lambda\), deriving an effect size is more challenging, as
\(\lambda\) does not have a sampling distribution from which the
standard error and confidence intervals may be obtained. Confidence
intervals are therefore generated for the values of \(\lambda\) that
intersect the log-likehihood profile for corresponding percentiles of
the \(\chi^2\) distribution used to compare the putative model to a null
model with \(\lambda = 0\) {[}\textbf{add ref: MLC thinks Boettiger
paper?}{]}. Thus, an effect size for \(\lambda\) may be found as:

\begin{align}
   \lvert Z_{\lambda} \rvert = \sqrt{\chi^2_{\hat{\lambda}}}
\end{align}

where, \(\hat{\lambda}\) is the maximized likelihood value of
\(\lambda\) and \(\chi^{2}_{\hat{\lambda}}\) is the likelihood ratio
statistic for the value.

Here we evaluate the ability of \(Z_{\lambda}\) and \(Z_{\kappa}\) to
characterize known levels of phylogenetic signal. Both \(Z_{\lambda}\)
and \(Z_{\kappa}\) are associated with input phylogenetic signal
(\(\lambda\)), indicating that both statistics capture the observed
signal (Fig. 3). However, effect sizes from \(\hat{\lambda}\) made
little sense, as they are more strongly associated with tree size than
they are with the actual phylogenetic signal in the data (Fig. 3). By
contrast, \(Z_{\kappa}\) is much more consistent accross tree sizes, and
increases more linearly with incrasing levels of phylogenetic signal.
Additionally, \(Z_{\kappa}\) exhibits a much stronger association with
phylogenetic signal strength as compared to tree size (Fig. 3), and its
standard deviation across input signal is more consistent. This implies
that similar levels of precision are found with \(Z_{\kappa}\) across
the range of input values. Thus between the two statistics,
\(Z_{\kappa}\) is a more reliable measure of the strength of
phylogenetic signal, and may be used to compare levels of phylogenetic
signal across datasets.

\begin{figure}
\centering
\includegraphics{new.fig.3.temp.png}
\caption{Response of effect sizes \(Z_\lambda\) and \(Z_\kappa\) to
increasing strength of Brownian motion.{}}
\end{figure}

\textbf{A test statistic (\(\hat{Z}_{12}\)) allows meaningful
comparisons across datasets.} To statistically compare the strength of
phylogenetic signal across datasets we propose a two-sample test
statistic (\(\hat{Z}_{12}\)). Based on statistical theory, a two-sample
test statistic may be calculated as:

\begin{align}
  \hat{Z}_{12}=\frac{\lvert{(\kappa_{1}-\hat\mu_{\kappa_1})-(\kappa_{2}-\hat\mu_{\kappa_2})}\rvert}{\sqrt{\hat\sigma^2_{\kappa_1}+\hat\sigma^2_{\kappa_2}}} = \frac{\lvert Z_{\kappa_1} - Z_{\kappa_2}\rvert}{\sqrt{2}}
\end{align}

where \(\kappa_1\), \(\kappa_2\), \(\hat\mu_{\kappa_1}\),
\(\hat\mu_{\kappa_2}\), \(\hat\sigma_{\kappa_1}\), and
\(\hat\sigma_{\kappa_2}\) are as defined above for equation 2. The right
side of the equation illustrates that if \(Z_\kappa\) has already been
calculated for two sampling distributions as in equation 2, the sampling
distributions have unit variance for each of the \(Z_\kappa\)
statistics. Estimates of significance of \(\hat{Z}_{12}\) may be
obtained from a standard normal distribution. Typically,
\(\hat{Z}_{12}\) is considered a two-tailed test, however directional
(one-tailed) tests may be specified should the empirical situation
require it (35, 37).

To demonstrate the utility of \(\hat{Z}_12\), we compared \(Z_{\kappa}\)
for two ecologically-relevant traits in plethodontid salamander (Fig.
4): surface area to volume ratios (SA:V) and relative body width
(\(\frac{BW}{SVL}\)) (39, 40). While both traits contained significant
phylogenetic signal, tests based on \(\hat{Z}_{12}\) revealed that the
degree of phylogenetic signal was significantly stronger in SA:V
(\(\hat{Z}_{12}=4.13\); \(P=0.000036\): Fig. 4). Biologically, this
observation may be interpreted by the fact that tropical species --
which form a monophyletic group within plethodontids -- display greater
variation in SA:V, which covaries with disparity in their climatic
niches (40). Thus, greater phylogenetic signal in SA:V is to be
expected.

\begin{figure}
\centering
\includegraphics{new.fig.4.png}
\caption{(A) Linear measures for relative body size, and regions of the
body used to estimate surface area to volume (SA:V) ratios. (B)
Permutation distributions of phylogenetic signal for SA:V and
\(\frac{BW}{SVL}\), with observed values shown as vertical bars. (C)
Effect sizes (\(Z_\kappa\)) for SA:V and \(\frac{BW}{SVL}\), with their
95\% confidence intervals (CI not standardized by \(\sqrt(n)\)).{}}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

It is common in comparative evolutionary studies to characterize the
phylogenetic signal in phenotypic traits to determine the extent to
which shared evolutionary history has generated trait covariation among
taxa. However, while numerous analytical approaches may be used to
quantify phylogenetic signal (11, 12, 14--16), methods that explicitly
measure the strength of phylogenetic signal, or facilitate comparisons
among datasets, have remained underdeveloped. We evaluated the precision
of one common measure, Pagel's \(\lambda\), and explored its efficacy
for characterizing the strength of phylogenetic signal in phenotypic
data. Using computer simulations, we found that \(\lambda\) behaves as a
Bernoulli random variable, with estimates that are increasingly skewed
at larger and smaller input levels of phylogenetic signal. Further, the
precision of \(\lambda\) in estimating actual levels of phylogenetic
signal varies with both tree size (see also ref. (23)) and input levels
of phylogentetic signal. From these findings we conclude that
\(\lambda\) is not a reliable indicator of the observed strength of
phylogenetic signal in phenoytpic datasets, and should not be used as an
effect size for comparing the degree of phylogenetic signal between
datasets.

As an alternative, we described a standardized effect size (\(Z\)) for
assessing the strength of phylogenetic signal. \(Z\) expresses the
magnitude of phylogenetic signal as a standard normal deviate, which is
easily interpretable as the strength of phylogenetic signal relative to
the mean. We applied this concept to both \(\lambda\) and \(\kappa\),
and found that \(Z_\kappa\) was a better estimate of the strength of
phylogenetic signal in phenotypic data. First, values of \(Z_\kappa\)
more accurately tracked known changes in the magnitude of phylogenetic
signal, as demonstrated by the linear relationship between \(Z_\kappa\)
and input signal. Additionally, the precision of \(Z_{\kappa}\) was more
consistent across the range of input levels of phylogenetic signal.
Thus, \(Z_\kappa\) is a more reliable measure of the relative strength
of phylogenetic signal, and places that effect on a common and
comparable scale. We therefore recommend that future studies interested
in evaluating the strength of phylogenetic signal incorporate
\(Z_\kappa\) as a statistical measure of this effect.

Next we proposed a two-sample test (\(\hat{Z}_{12}\)), which provides a
formal statistical procedure for determining whether the strength of
phylogenetic signal is greater in one phenotypic trait as compared to
another. Prior studies have summarized patterns of variation in
phylogenetic signal across datasets using summary test values, such as
\(\kappa\) (12). However, because \(\kappa\) does not scale linearly
with input levels of phylogenetic signal (Fig. 2), and its variance
increases with increasing strength of phylogenetic signal (18, 20), it
should not be considered an effect size that measures the strength of
phylogenetic signal on a common scale. By contrast, standardizing
\(\kappa\) to \(Z_\kappa\) via equation 2 alleviates these concerns, and
facilitates formal statistical comparisons of the strength of signal
across datasets. Thus when viewed from this perspective, the approach
developed here aligns well with other statistical approaches such as
meta-analysis (31, 41, 42), where summary statistics across datasets are
converted to standardized effect sizes for subsequent ``higher order''
statistical summaries or comparisons. As such, our approach enables
evolutionary biologists to quantitatively examine the relative strength
of phylogenetic signal across a wide range of phenotypic traits, and
thus opens the door for future discoveries that inform on how phenotypic
diversity accumulates in macroevolutionary time across the tree of life.

One important advantage of the approach advocated here is that the
resulting effect sizes (\(Z_\kappa\)) are dimensionless, as the units of
measurement cancel out during the calculation of \(Z\) (43). Thus,
\(Z_\kappa\) represents the strength of phylogenetic signal on a common
and comparable scale -- measured in standard deviations -- regardless of
the initial units and original scale of the phenotypic variables under
investigation. This means that the strength of phylogenetic signal may
be compared across datasets for continuous phenotypic traits measured in
different units and scale, because those units have been standardized
through their conversion to \(Z_\kappa\). For example, our approach
could be utilized to determine whether the strength of phylogenetic
signal (say, in response to ecological differentiation) is stronger in
morphological traits (linear traits: \(mm\)), physiological traits
(metabolic rate: \(\frac{O^2}{min}\)), or behavioral traits (aggression:
\(\frac{\#{displays}}{second}\)). In fact, our empirical example
provided just such a comparison, as SA:V is represented in \(mm^{-1}\)
while relative body size is a unitless ratio (\(\frac{BW}{SVL}\)).
Additionally, our method is capable of comparing the strength of
phylogenetic signal in traits of different dimensionality, as estimates
of phylogenetic signal using \(\kappa\) have been generalized for
multivariate data (16). Furthermore, tests based on \(\hat{Z}_{12}\) may
be utilized for comparing the strength of phylogenetic signal among
datasets containing a different number of species, and even for
phenotypes obtained from species in different lineages, because their
phylogenetic non-independence and observed variation are taken into
account in the generation of the empirical sampling distribution via
permutation.

This study is not the first to compare \(\lambda\) and \(\kappa\) for
their ability as statistics to measure phylogenetic signal. Our results
for \(\lambda\) and \(\kappa\) values are consistent with those found in
the simulations performed by Münkemüller et al.~(18), but that study
investigated type I error rates and statistical power, finding that
\(\lambda\) performed better in both regards, irrespective of species
number in trees. Although not the central focus of their study, the same
tendency for variable \(\lambda\) and consistent \(\kappa\) at
intermediate phylogenetic signal strengths was observed (Fig. 2 of ref.
(18)). Recent work by Molina-Venegas and Rodríguez (21) found that
\(\kappa\) but not \(\lambda\) tended to inflate the estimate of
phylogenetic signal, leading to moderate type I and type II biases, if
polytomic chronograms were used. Their work more thoroughly addressed
previous observations of inflated \(\kappa\) for incompletely resolved
phylogenetic trees (18, 44). An interesting question is whether an
inflated \(\kappa\) value leads to an inflated \(Z_\kappa\) or does a
tendency of a particular tree to inflate estimates of \(\kappa\) also
inflate the values in random permutations of a test, in which case
\(Z_\kappa\) is robust to polytomies? We repeated the analyses in Figs.
1 \& 2, adjusting trees to have 50\% collapsed nodes, per the technique
of Molina-Venegas and Rodríguez (21), and found results were consistent
(Supporting Information). This confirms that any tendency of
incompletely resolved trees to inflate \(\kappa\) as a descriptive
statistic does not inflate \(Z_\kappa\) as an effect size. Furthermore,
because comparison of effect sizes in a test is a comparison of
locations of observed values in their sampling distributions, which
would shift concomitantly because of this tendency, the \(Z_{12}\) test
statistic in equation 4 appears to be robust in spite of unresolved
trees.

Phylogenetic signal can be thought of as both an attribute to be
measured in the data and a parameter that can be tuned to account for
the phylogenetic non-independence among observations, for analysis of
the data. As such, \(\lambda\) is appealing, as a statistic that
potentially fulfills both roles. However, the inability to estimate
phylogenetic signal with \(\lambda\) for data simulated with known
phylogenetic signal is troublesome, and we recommend evolutionary
biologists refrain from viewing it as a statistic to describe the amount
of phylogenetic signal in the data. Interestingly, \(\kappa\) -- when
standardized to an effect size \(Z_\kappa\) -- is a better statistic for
measuring the amount of phylogenetic signal in data simulated with
respect to known levels of \(\lambda\). Although \(\lambda\) might be
viewed as an important parameter for modifying the the conditional
estimation of linear model coefficients with respect to phylogeny, it is
neither a statistic that has meaningful comparative value as a measure
of phylogenetic signal nor a statistic that lends itself well to
reliable calculation of a test statistic. By contrast, \(\kappa\) has
been shown here to be a reliable statistic, but only when standardized
by the mean and standard deviation of its empirical sampling
distribution (i.e., when converted to the effect size, \(Z_\kappa\)).
Because one has control over the number of permutations used in
analysis, one can be assured with many permutations that the empirical
sampling distribution is representative of true probability
distributions (10). Given the greater consistency in estimates of
\(Z_\kappa\) across tree sizes and input signal, it is difficult to
imagine a hypothesis test that can improve equation 4 for efficiently
comparing phylogenetic signal for different traits, different trees, or
a combination of both.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\textbf{Simulations.} Simulations were conducted by generating
pure-birth phylogenies at each of six different tree sizes
(\(n=2^5, 2^6, \cdots, 2^{10}\)), and with differing levels of
phylognetic signal (\(\lambda=0.0, 0.5, \cdots, 1.0\)). We generated 100
random trees for each intersection of tree size and \(\lambda\). For
each \(\lambda\) within each tree size, continuous traits were then
simulated on each phylogeny under a BM model of evolution. For each set
of 100 trees we measured the mean values of \(\hat{\lambda}\) and
\(\kappa\), their standard deviation, and calculated the Shapiro-Wilk
\(W\) statistic as a departure from normality (symmetry). For the
latter, a value of \(1.0\) indicates normally distributed values, while
departures from \(1.0\) indicate skewness. Simulations were then
repeated for both bananced and pectinate trees, which yielded
qualitatively similar results (see Supporting Information). Trees
containing polytomies, and an evaluation of \(\hat{\lambda}\) from
models of linear regression and phylogenetic ANOVA, were also
investigated, and results were qualitatively similar to those reported
above (see Supporting Information).

\textbf{Empirical Data.} Surface area to volume ratios (SA:V) and
relative body width (\(\frac{BW}{SVL}\)) measures were obtained from as
species means from individuals of 305 species, from which species means
were obtained (39, 40). A time-dated molecular phylogeny for the group
(45) was pruned to match the species in the phenotypic dataset. The
phylogenetic signal in each trait was then characterized using
\(\kappa\), which was converted to its effect size (\(Z_\kappa\)) using
\texttt{geomorph} 3.3.1 (46, 47), and routines by the authors
(\textbf{to be incorporated in \texttt{geomorph} upon manuscript
acceptance}).

\showmatmethods
\showacknow
\pnasbreak

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Felsenstein1985}{}%
1. Felsenstein J (1985) Phylogenies and the comparative method.
\emph{American Naturalist} 125(1):1--15.

\leavevmode\hypertarget{ref-HarveyPagel1991}{}%
2. Harvey PH, Pagel MD (1991) \emph{The comparative method in
evolutionary biology} (Oxford University Press, Oxford).

\leavevmode\hypertarget{ref-Grafen1989}{}%
3. Grafen A (1989) The phylogenetic regression. \emph{Philosophical
Transactions of the Royal Society of London B, Biological Sciences}
326:119--157.

\leavevmode\hypertarget{ref-GarlandIves2000}{}%
4. Garland TJ, Ives AR (2000) Using the past to predict the present:
Confidence intervals for regression equations in phylogenetic
comparative methods. \emph{American Naturalist} 155:346--364.

\leavevmode\hypertarget{ref-Rohlf2001}{}%
5. Rohlf FJ (2001) Comparative methods for the analysis of continuous
variables: Geometric interpretations. \emph{Evolution} 55:2143--2160.

\leavevmode\hypertarget{ref-MartinsHansen1997}{}%
6. Martins EP, Hansen TF (1997) Phylogenies and the comparative method:
A general approach to incorporating phylogenetic information into the
analysis of interspecific data. \emph{American Naturalist} 149:646--667.

\leavevmode\hypertarget{ref-OMeara_et_al2006}{}%
7. O'Meara BC, Ane C, Sanderson MJ, Wainwright PC (2006) Testing for
different rates of continuous trait evolution using likelihood.
\emph{Evolution} 60:922--933.

\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
8. Beaulieu JM, Jhwueng DC, Boettiger C, O'Meara BC (2012) Modeling
stabilizing selection: Expanding the ornstein-uhlenbeck model of
adaptive evolution. \emph{Evolution} 66:2369--2383.

\leavevmode\hypertarget{ref-Adams2014b}{}%
9. Adams DC (2014) A method for assessing phylogenetic least squares
models for shape and other high-dimensional multivariate data.
\emph{Evolution} 68:2675--2688.

\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
10. Adams DC, Collyer ML (2018) Phylogenetic anova: Group-clade
aggregation, biological challenges, and a refined permutation procedure.
\emph{Evolution} 72(6):1204--1215.

\leavevmode\hypertarget{ref-Pagel1999}{}%
11. Pagel MD (1999) Inferring the historical patterns of biological
evolution. \emph{Nature} 401:877--884.

\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
12. Blomberg SP, Garland T, Ives AR (2003) Testing for phylogenetic
signal in comparative data: Behavioral traits are more labile.
\emph{Evolution} 57:717--745.

\leavevmode\hypertarget{ref-Revell_et_al2008}{}%
13. Revell LJ, Harmon LJ, Collar DC (2008) Phylogenetic signal,
evolutionary process, and rate. \emph{Systematic Biology} 57:591--601.

\leavevmode\hypertarget{ref-Abouheif1999}{}%
14. Abouheif E (1999) A method for testing the assumption of
phylogenetic independence in comparative data. \emph{Evolutionary
Ecology Research} 1:895--909.

\leavevmode\hypertarget{ref-Gittleman1990}{}%
15. Gittleman JL, Kot M (1990) Adaptation: Statistics and a null model
for estimating phylogenetic effects. \emph{Systematic Zoology}
39(3):227--241.

\leavevmode\hypertarget{ref-Adams2014a}{}%
16. Adams DC (2014) A generalized Kappa statistic for estimating
phylogenetic signal from shape and other high-dimensional dultivariate
data. \emph{Systematic Biology} 63:685--697.

\leavevmode\hypertarget{ref-Klingenberg2010}{}%
17. Klingenberg CP, Gidaszewski NA (2010) Testing and quantifying
phylogenetic signals and homoplasy in morphometric data.
\emph{Systematic biology} 59(3):245--261.

\leavevmode\hypertarget{ref-Munkemuller_et_al2012}{}%
18. Münkemüller T, et al. (2012) How to measure and test phylogenetic
signal. \emph{Methods in Ecology and Evolution} 3:743--756.

\leavevmode\hypertarget{ref-Pavoine2012}{}%
19. Pavoine S, Ricotta C (2012) Testing for phylogenetic signal in
biological traits: The ubiquity of cross-product statistics.
\emph{Evolution: International Journal of Organic Evolution}
67(3):828--840.

\leavevmode\hypertarget{ref-DinizFilho2012}{}%
20. Diniz-Filho JAF, Santos T, Rangel TF, Bini LM (2012) A comparison of
metrics for estimating phylogenetic signal under alternative
evolutionary models. \emph{Genetics and Molecular Biology}
35(3):673--679.

\leavevmode\hypertarget{ref-MolinaVenegas2017}{}%
21. Molina-Venegas R, Rodríguez MA (2017) Revisiting phylogenetic
signal; strong or negligible impacts of polytomies and branch length
information? \emph{BMC evolutionary biology} 17(1):53.

\leavevmode\hypertarget{ref-Revell2010}{}%
22. Revell LJ (2010) Phylogenetic signal and linear regression on
species data. \emph{Methods in Ecology and Evolution} 1:319--329.

\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
23. Boettiger C, Coop G, Ralph P (2012) Is your phylogeny informative?
Measuring the power of comparative methods. \emph{Evolution}
67:2240--2251.

\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
24. Freckleton RP, Harvey PH, Pagel M (2002) Phylogenetic analysis and
comparative data: A test and review of evidence. \emph{American
Naturalist} 160:712--726.

\leavevmode\hypertarget{ref-Cooper2010}{}%
25. Cooper N, Jetz W, Freckleton RP (2010) Phylogenetic comparative
approaches for studying niche conservatism. \emph{Journal of
Evolutionary Biology} 23(12):2529--2539.

\leavevmode\hypertarget{ref-Bose2019}{}%
26. Bose R, Ramesh BR, Pélissier R, Munoz F (2019) Phylogenetic
diversity in the western ghats biodiversity hotspot reflects
environmental filtering and past niche diversification of trees.
\emph{Journal of Biogeography} 46(1):145--157.

\leavevmode\hypertarget{ref-Vandelook2019}{}%
27. Vandelook F, et al. (2019) Nectar traits differ between pollination
syndromes in balsaminaceae. \emph{Annals of Botany} 124(2):269--279.

\leavevmode\hypertarget{ref-DeMeester2019}{}%
28. De Meester G, Huyghe K, Van Damme R (2019) Brain size, ecology and
sociality: A reptilian perspective. \emph{Biological Journal of the
Linnean Society} 126(3):381--391.

\leavevmode\hypertarget{ref-Pintanel2019}{}%
29. Pintanel P, Tejedo M, Ron SR, Llorente GA, Merino-Viteri A (2019)
Elevational and microclimatic drivers of thermal tolerance in andean
pristimantis frogs. \emph{Journal of Biogeography} 46(8):1664--1675.

\leavevmode\hypertarget{ref-Su2019}{}%
30. Su G, Villéger S, Brosse S (2019) Morphological diversity of
freshwater fishes differs between realms, but morphologically extreme
species are widespread. \emph{Global ecology and biogeography}
28(2):211--221.

\leavevmode\hypertarget{ref-Glass1976}{}%
31. Glass GV (1976) Primary, secondary, and meta-analysis of research.
\emph{Educational Researcher} 5:3--8.

\leavevmode\hypertarget{ref-Cohen1988}{}%
32. Cohen J (1988) \emph{Statistical power analysis for the behavioral
sciences} (Routledge).

\leavevmode\hypertarget{ref-Rosenthal1994}{}%
33. Rosenthal R (1994) The handbook of research synthesis. ed Cooper LV
H Hedges (Russell Sage Foundation), pp 231--244.

\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
34. Collyer ML, Sekora DJ, Adams DC (2015) A method for analysis of
phenotypic change for phenotypes described by high-dimensional data.
\emph{Heredity} 115:357--365.

\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
35. Adams DC, Collyer ML (2016) On the comparison of the strength of
morphological integration across morphometric datasets. \emph{Evolution}
70:2623--2631.

\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
36. Collyer ML, Adams DC (2018) RRPP: An r package for fitting linear
models to high-dimensional data using residual randomization.
\emph{Methods in Ecology and Evolution} 9:1772--1779.

\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
37. Adams DC, Collyer ML (2019) Comparing the strength of modular
signal, and evaluating alternative modular hypotheses, using covariance
ratio effect sizes with morphometric data. \emph{Evolution}
73(12):2352--2367.

\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
38. Adams DC, Collyer ML (2019) Phylogenetic comparative methods and the
evolution of multivariate phenotypes. \emph{Annual Review of Ecology,
Evolution, and Systematics} 50:405--425.

\leavevmode\hypertarget{ref-Baken2019}{}%
39. Baken EK, Adams DC (2019) Macroevolution of arboreality in
salamanders. \emph{Ecology and Evolution} 9(12):7005--7016.

\leavevmode\hypertarget{ref-Baken2020}{}%
40. Baken EK, Mellenthin LE, Adams DC (2020) Macroevolution of
desiccation‐related morphology in plethodontid salamanders as inferred
from a novel surface area to volume ratio estimation approach.
\emph{Evolution} 74:476--486.

\leavevmode\hypertarget{ref-HedgesOlkin1985}{}%
41. Hedges L. V., Olkin I (1985) \emph{Statistical methods for
meta-analysis} (Elsevier).

\leavevmode\hypertarget{ref-Arnqvist1995}{}%
42. Arnqvist G., Wooster D (1995) Meta-analysis: Synthesizing research
findings in ecology and evolution. \emph{Trends in Ecology and
Evolution} 10:236--240.

\leavevmode\hypertarget{ref-SokalRohlf2012}{}%
43. Sokal R. R., Rohlf FJ (2012) \emph{Biometry} (W.H. Freeman \& Co.,
San Francisco). 4th Ed.

\leavevmode\hypertarget{ref-Davies2012}{}%
44. Davies TJ, Kraft NJ, Salamin N, Wolkovich EM (2012) Incompletely
resolved phylogenetic trees inflate estimates of phylogenetic
conservatism. \emph{Ecology} 93(2):242--247.

\leavevmode\hypertarget{ref-Bonett2017}{}%
45. Bonett RM, Blair AL (2017) Evidence for complex life cycle
constraints on salamander body form diversification. \emph{Proceedings
of the National Academy of Sciences, USA} 114:9936--9941.

\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
46. Adams DC, Otárola-Castillo E (2013) Geomorph: An r package for the
collection and analysis of geometric morphometric shape data.
\emph{Methods in Ecology and Evolution} 4:393--399.

\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
47. Adams DC, Collyer ML, Kaliontzopoulou A (2020) Geomorph: Software
for geometric morphometric analyses. R package version 3.3.1. Available
at: \url{https://cran.r-project.org/package=geomorph}.

% Bibliography
% \bibliography{pnas-sample}

\end{document}